### 2023-06-06 manuscript_soy.tex
Initial Overleaf Import

 \documentclass[9pt, onecolumn,twoside]{gsajnl}
% how i want it for inport into word: \documentclass[9pt, onecolumn,twoside]{gsajnl}
% how i want it for pdf: \documentclass[9pt,twocolumn,twoside,lineno]{gsajnl}
% Use the documentclass option 'lineno' to view line numbers

\usepackage{epstopdf}

\articletype{gs} % article type
% {inv} Investigation
% {gs} Genomic Selection
% {goi} Genetics of Immunity
% {gos} Genetics of Sex
% {mp} Multiparental Populations

\runningtitle{Soy manuscript} % For use in the footer
\runningauthor{Connelly \textit{et al.}}

\title{Working title SoyAdapt}

\author[1,$\dagger$,$\ast$]{Josephine Estelle Ananda Connelly}
\author[2]{Guillaume Paul Ramstein}
\author[3]{Wolf L Eiserhardt}
\author[4]{Torben Asp}
\author[5]{Josephine Estelle Ananda Connelly}

\affil[1]{Centre for Quantitative Genetics and Genomics, Faculty of Technical Sciences, Aarhus University (DK).}
\affil[2]{Department of Biology, Faculty of Natural Sciences, Aarhus University, (DK).}
\affil[3]{Royal Botanic Gardens, Kew, (UK).}

\correspondingauthoraffiliation[$\ast$]{Corresponding author: AU, Connelly.}

\begin{abstract}
Soybean (Glycine max L. Merrill) is the worlds leading  oilseed crop used as a  primary source of vegetable oil for human consumption and in proteinmeal for animal feed.
\end{abstract}

\keywords{Soy; population genetics; Diversity; Selection; more keywords}

\dates{\rec{01 09, 2023} \acc{01 09, 2023}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
%\slugnote
%\firstpagefootnote
\vspace{-13pt}% Only used for adjusting extra space in the left column of the first page

\section{Abstract}

Soybean, belonging to the Fabaceae family and scientifically known as \textit{Glycine max} (L. Merrill), is globally recognized as the primary legume crop due to its high protein and oil content. However, its genetic diversity has decreased over time due to domestication and intensive artificial selection, driving the need to look elsewhere to increase germplasm utilization for abiotic adaptation \citep{hyten06, gizlice96}.  Relevant accessions with the potential for abiotic adaptation to cold tolerance have been identified [@haupt20]. Additionally 160 Swedish accessions originating from a Swedish breed for the adaptation to the cool climate of north western Europe and stored in the NordGen gene bank have been unearthed [@holmberg1973]. We will interpret the plant genetic resources available, so as to utilise the genetic potential. To facilitate targeted selection of promising accessions from soybeans vast germplasm collections,  and unveiling the genetic data stored in the NordGen seed banks. findings are: ...

\section{Introduction}

\textit{what we already know:} Several severe genetic bottlenecks occurred in soybean. Compared to the wild species, the genetic diversity was halved, resulting in a loss of 81\% of rare alleles (Hyten et al., 2006). Additionally, two major bottlenecks occurred during the development of North American modern cultivars, where only a few landraces were used. And as a result of intensive soybean breeding over the past 75 years. Elite cultivars have emerged, but they are derived from only about 19 landraces. Consequently, the North American breeding pools now retain only 72\% of genome diversity and have lost 79\% of rare alleles found in diverse landraces (Gizlice et al., 1996).

\textit{soybean needs to be grown in the north}:  This argument but then for spread of growth area! "Modern U.S. soybean breeding has led to a yield increase of 29 kg ha/−1 yr/−1 (Rincker et al., 2014)" 
 ""Better understanding the genomic basis behind this improvement may provide indicators for further soybean improvement and adaptation. "  from: Sequencing the USDA core soybean collection reveals gene loss during domestication and breeding

\textit{the swedish soybeans (goals and scope)}

\textit{mention approaches}

\textit{In this study we}
The goal is to help precipitate the utilization of these soybean genetic resources.

\section{Materials and methods}
\label{sec:materials:methods}

what is the data: 160 soybean accessions were obtained from the Nordgen genebank, 155 of these accessions are of Swedish origin. These accessions are thought to originate from a Swedish breeding program running from the 1840s to the 1970s which used material consisting of a mixture of elite cultivars of the time and Japanese landraces, breeding for the adaptation to the cool climate of north western Europe \cite{holmberg1973}.

The Core collection used here is a subset of the from a large soybean germplasm collection, the  USDA genebank accessions consisting of 415 of the USDA soy germplasm accessions. This Core collection has been selected with a focus on adaptation to high-latitude cold regions using environmental data from phenotypic trials in Germany, and comparing Donor opulation of Environments (DPE) in Asia and the Target Population of Environments (TPE) in Central Europe. From the 3663 accessions two diverse core collections of 183 and 366 accessions were created. These 514 diversity panels are used here due to that they are likely preadapted to cultivation in Central Europe, while simultaneously conserving a high level of genetic diversity .

USDA germplasm of soybean consists of ... soybase.org.  Founders 10.

wgs
What was done: The accessions from the Nordgen genebank and the Core collection accessions were whole genome sequenced. Four of the Core collection seeds didn't germinate in time for the sequencing, resulting in accessions sequenced 574. Of the Nordgen accessions, Four were of Latvian origin and therefore excluded. 
GATK: 
The reads were aligned to the Williams 82 2a reference genome ( (Schmutz et al. 2010).) adapted from GATK Best Practices, keeping MQ30 and biallelic sites and applied a MAF filter of 0.01 in order to remove rare variants from the data, reducing the number of SNPs from 17,648,123 to 10,000,122. 
For a detailed account of programs and commands used see supplementary material: \nameref{sec:supplementary:material} 
From when i received it:
A further three accessions were removed due to missing metadata and a single accessions with missing data > 5\% using BCFtools \cite{bcftools}, resulting in 153 nordgen accessions and 409 CCA accessions and a SNP count of 8,533,444.

SNP data
The snp data is a high-density genotyping array for soybean with 47,337 single nucleotide polymorphisms (SNPs) was develop from soybean (Glycine max L. Merr.). The SoySNP50K iSelect BeadChip has been used to genotype the USDA Soybean Germplasm Collection and is available for use from soybase
 An intersect was made of the whole genome sequenced
35486

For a detailed account of programs and commands used see supplementary material: \nameref{sec:supplementary:material}

H1
Q 
To infer population stratification of the Swedish accessions from the NordGen gene bank and the Core collection and of the 12 accessions assumed to be founders of the Swedish breeding program.

Method
pca
Principal component analysis were applied using the package SnpRelate in R (A Parallel Computing Toolset for Relatedness and Principal Component Analysis of SNP Data).

IBS Matrix: The package SNPRelate in R was used to compute an identity-by-state (IBS) matrix that calculates the proportion that two randomly selected reads that contain a certain SNP locus are the same or different between two individuals. The resulting pairwise IBS matrix was used to generate a cluster dendogram using the function \textit{snpgdsHCluster: Hierarchical cluster analysis In SNPRelate }in R.

 snpgdsHCluster: Hierarchical cluster analysis
In SNPRelate: Parallel Computing Toolset for Relatedness and Principal Component Analysis of SNP Data

realisation that 2 of the swedish were likely not a part of the swedish breeding program and excluded 
see Supplimentary material for accession data. 
Fst of the swedish vs the cca

H2 
Q diversity

Method
LD
pi and

H3
Q

Method

\subsection{Statistical analysis}

Indicate what statistical analysis has been performed.

\section{Results and Discussion}
\subsection{population structure and genetic differentiation}

H1: 
To better understand the population relationships of the accessions received from the Nordic gene bank and how they are related to each other, and the rest of the soybean germplasm, they are compared to a core collection of 409 soybean from the USDA genebank selected from the soybean (Glycine max) germplasm for Central European breeding.  The Core collection used here is a subset of the from a large soybean germplasm collection, the  USDA genebank accessions, that are likely preadapted to cultivation in Central Europe, while simultaneously conserving a high level of genetic diversity (CCA paper). Additionally accessions were grouped into categories based on known breeding history, So the suspected Founders of the Swedish breeding program are grouped together. Here an intersect of the whole genome sequences the Core Collection accessions, the Swedish Nordgen accessions and the 50kSNP data available of the ten accessions identified as possible genetic Founders to the Swedish breeding program is used.

To understand the population differentiation/ diversity, a Principal Component Analysis (PCA) and the genome-wide Identity By State (IBS) pairwise distance matrix were applied. The resulting PCA shows the Swedish accessions as a sub group seen in green in (Figure \ref{fig:pca}) . A closer look at the Nordgen Swedish accessions reviled two accessions that were not part of the particular breeding program as seen in (Figure X showing the SBPA PCA with circles around the two accessions not from the breeding program)  and also thereafter confirmed by the genebank passport data. (Or do i simply remove these from the analysis and only mention that they are excluded up in the methods when describing the data)

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{plot_PCA_origin.pdf}
\caption{This PCA shows the CCA in black, the SBPA in green and the thought to be founders in red.}%
\label{fig:pca}
\end{figure}

 and the IBS (Figure  \ref{fig:dendo}) shows . 
 IBS numbers in a table showing ibs within sbpa and out?

The genetic differentiation measured by FST

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{rainbow.pdf}
\caption{Hierarchical cluster dendrogram based on pairwise identity-by-state (IBS) values from WGS data for all samples. }%
\label{fig:dendo}
\end{figure}

\subsection{Genetic diversity}

MAF?
LD
pi 
theta

\subsection{Genome wide selection signatures} 
gws.

\section{Discussion}

The Swedish soybeans are are mainly a snapshot of a breeding program stopped around 1978 and all lines or recent crosses from this program were then set into the

\section{Additional guidelines}

\subsection{Numbers} In the text, write out numbers nine or less except as part of a date, a fraction or decimal, a percentage, or a unit of measurement. Use Arabic numbers for those larger than nine, except as the first word of a sentence; however, try to avoid starting a sentence with such a number.

\subsection{Units} Use abbreviations of the customary units of measurement only when they are preceded by a number: "3 min" but "several minutes". Write "percent" as one word, except when used with a number: "several percent" but "75\%." To indicate temperature in centigrade, use ° (for example, 37°); include a letter after the degree symbol only when some other scale is intended (for example, 45°K).

\subsection{Nomenclature and italicization} Italicize names of organisms even when  when the species is not indicated.  Italicize the first three letters of the names of restriction enzyme cleavage sites, as in HindIII. Write the names of strains in roman except when incorporating specific genotypic designations. Italicize genotype names and symbols, including all components of alleles, but not when the name of a gene is the same as the name of an enzyme. Do not use "+" to indicate wild type. Carefully distinguish between genotype (italicized) and phenotype (not italicized) in both the writing and the symbolism.

\subsection{Cross references}
Use the \verb|\nameref| command with the \verb|\label| command to insert cross-references to section headings. For example, a \verb|\label| has been defined in the section \nameref{sec:materials:methods}.

\section{In-text citations}

Add citations using the \verb|\citep{}| command, for example \citep{neher2013genealogies} or for multiple citations, \citep{neher2013genealogies, rodelsperger2014characterization,Falush16}

\section{Examples of article components}
\label{sec:examples}

The sections below show examples of different header levels, which you can use in the primary sections of the manuscript (Results, Discussion, etc.) to organize your content.

\section{First level section header}

Use this level to group two or more closely related headings in a long article.

\subsection{Second level section header}

Second level section text.

\subsubsection{Third level section header:}

Third level section text. These headings may be numbered, but only when the numbers must be cited in the text.

\section{Figures and tables}

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

\subsection{PCA figure}

Figure \ref{fig:pca} shows an example figure.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{plot_PCA_origin.pdf}
\caption{This PCA shows the CCA in black, the SBPA in green and the thought to be founders in red.}%
\label{fig:pca}
\end{figure}

%\begin{figure}[htbp]
%\centering
%\includegraphics[width=\linewidth]{example-figure}
%%\caption{Example movie (the figure file above is used as a placeholder for this example). \textit{GENETICS} supports video and movie files that can be linked from any portion of the article - including the abstract. Acceptable formats include .asf, avi, .wav, and all types of Windows Media files.
%%}%
%%\label{video:spectrum}
%\end{figure}

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

\subsection{Dendogram figure}

Figure \ref{fig:dendo} shows an example figure.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{rainbow.pdf}
\caption{Hierarchical cluster dendrogram based on pairwise identity-by-state (IBS) values from SNP data for all samples. describe dendogram}%
\label{fig:pca}
\end{figure}

%\begin{figure}[htbp]
%\centering
%\includegraphics[width=\linewidth]{example-figure}
%%\caption{Example movie (the figure file above is used as a placeholder for this example). \textit{GENETICS} supports video and movie files that can be linked from any portion of the article - including the abstract. Acceptable formats include .asf, avi, .wav, and all types of Windows Media files.
%%}%
%%\label{video:spectrum}
%\end{figure}

Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.

\subsection{Sample figure}

Figure \ref{fig:spectrum} shows an example figure.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{example-figure}
\caption{Example figure from \url{10.1534/genetics.114.173807}. Please include your figures in the manuscript for the review process. You can upload figures to Overleaf via the Project menu. Images of photographs or paintings can be provided as raster images. Common examples of raster images are .tif/.tiff, .raw, .gif, and .bmp file types. The resolution of raster files is measured by the number of dots or pixels in a given area, referred to as “dpi” or “ppi.”
\begin{itemize}
\item  minimum resolution required for printed images or pictures: 350dpi
\item  minimum resolution for printed line art: 600dpi (complex or finely drawn line art should be 1200dpi)
\item minimum resolution for electronic images (i.e., for on-screen viewing): 72dpi
\protect\end{itemize}
Images of maps, charts, graphs, and diagrams are best rendered digitally as geometric forms called vector graphics. Common file types are .eps, .ai, and .pdf. Vector images use mathematical relationships between points and the lines connecting them to describe an image. These file types do not use pixels; therefore resolution does not apply to vector images.
Label multiple figure parts with A, B, etc. in bolded. Legends should start with a brief title and should be a self-contained description of the content of the figure that provides enough detail to fully understand the data presented. All conventional symbols used to indicate figure data points are available for typesetting; unconventional symbols should not be used. Italicize all mathematical variables (both in the figure legend and figure) , genotypes, and additional symbols that are normally italicized.}%
\label{fig:spectrum}
\end{figure}

%\begin{figure}[htbp]
%\centering
%\includegraphics[width=\linewidth]{example-figure}
%%\caption{Example movie (the figure file above is used as a placeholder for this example). \textit{GENETICS} supports video and movie files that can be linked from any portion of the article - including the abstract. Acceptable formats include .asf, avi, .wav, and all types of Windows Media files.
%%}%
%%\label{video:spectrum}
%\end{figure}

\subsection{Sample table}

Table \ref{tab:shape-functions} shows an example table. Avoid shading, color type, line drawings, graphics, or other illustrations within tables. Use tables for data only; present drawings, graphics, and illustrations as separate figures. Histograms should not be used to present data that can be captured easily in text or small tables, as they take up much more space.

Tables numbers are given in Arabic numerals. Tables should not be numbered 1A, 1B, etc., but if necessary, interior parts of the table can be labeled A, B, etc. for easy reference in the text.

\begin{table*}[p]
\centering
\caption{Students and their grades}
\begin{tableminipage}{\textwidth}
\begin{tabularx}{\textwidth}{@{}XXXX@{}}
\hline
{\bf Student} & {\bf Grade}\footnote{This is an example of a footnote in a table. Lowercase, superscript italic letters (a, b, c, etc.) are used by default. You can also use *, **, and *** to indicate conventional levels of statistical significance, explained below the table.} & {\bf Rank} & {\bf Notes} \\
\hline
Alice & 82\% & 1 & Performed very well.\\
Bob & 65\% & 3 & Not up to his usual standard.\\
Charlie & 73\% & 2 & A good attempt.\\
\hline
\end{tabularx}
  \label{tab:shape-functions}
\end{tableminipage}
\end{table*}

\section{Sample equation}

Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
\begin{equation}
S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
      = \frac{1}{n}\sum_{i}^{n} X_i
\label{eq:refname1}
\end{equation}
denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.

\section{Supplementary Material}
\label{sec:supplementary:material}
Supplementary Material Methods (S Methods)

\section{Data availability}

For example: Strains and plasmids are available upon request. File S1 contains detailed descriptions of all supplemental files. File S2 contains SNP ID numbers and locations. File S3 contains genotypes for each individual. Sequence data are available at GenBank and the accession numbers are listed in File S3. Gene expression data are available at GEO with the accession number: GDS1234. Code used to generate the simulated data can be found at \url{https://github.com/JosephineConnelly/soyadapt_data_analysis}.

\section{Acknowledgments}
Acknowledgments should be included here.

\section{Funding}
Funding, including Funder Names and Grant numbers should be included here.

\section{Conflicts of interest}
There are  no known conflicts of interest.

\bibliography{bibliography}

\end{document}

### 2023-06-07 manuscript_soy.tex
7 june backup

@@
 \author[4]{Torben Asp}
-\author[5]{Josephine Estelle Ananda Connelly}
 
-\affil[1]{Centre for Quantitative Genetics and Genomics, Faculty of Technical Sciences, Aarhus University (DK).}
-\affil[2]{Department of Biology, Faculty of Natural Sciences, Aarhus University, (DK).}
+\affil[1,2,4]{Centre for Quantitative Genetics and Genomics, Faculty of Technical Sciences, Aarhus University (DK).}
+\affil[1,3]{Department of Biology, Faculty of Natural Sciences, Aarhus University, (DK).}
 \affil[3]{Royal Botanic Gardens, Kew, (UK).}
 
-\correspondingauthoraffiliation[$\ast$]{Corresponding author: AU, Connelly.}
+\correspondingauthoraffiliation[$\ast$]{Corresponding author: AU, JosephineConnelly@qgg.uni.dk.}
 
@@
 Soybean, belonging to the Fabaceae family and scientifically known as \textit{Glycine max} (L.
-Merrill), is globally recognized as the primary legume crop due to its high protein and oil content.
-However, its genetic diversity has decreased over time due to domestication and intensive artificial selection, driving the need to look elsewhere to increase germplasm utilization for abiotic adaptation \citep{hyten06, gizlice96}.
+Merrill), is globally recognised as the primary legume crop due to its high protein and oil content.
+However, its genetic diversity has decreased over time due to domestication and intensive artificial selection, driving the need to look elsewhere to increase germplasm utilisation for abiotic adaptation \citep{hyten06, gizlice96}.
 Relevant accessions with the potential for abiotic adaptation to cold tolerance have been identified [@haupt20].
@@
 soybean breeding has led to a yield increase of 29 kg ha/−1 yr/−1 (Rincker et al., 2014)"
- ""Better understanding the genomic basis behind this improvement may provide indicators for further soybean improvement and adaptation.
-"  from: Sequencing the USDA core soybean collection reveals gene loss during domestication and breeding
+ "Better understanding the genomic basis behind this improvement may provide indicators for further soybean improvement and adaptation.
+"  from: Sequencing the USDA core soybean collection reveals gene loss during domestication and breeding"
 
@@
 \textit{In this study we}
-The goal is to help precipitate the utilization of these soybean genetic resources.
+The goal is to help precipitate the utilisation of these soybean genetic resources.
 
@@
 wgs
-What was done: The accessions from the Nordgen genebank and the Core collection accessions were whole genome sequenced.
+\textit{What was done:} The accessions from the Nordgen genebank and the Core collection accessions were whole genome sequenced.
 Four of the Core collection seeds didn't germinate in time for the sequencing, resulting in accessions sequenced 574.
 Of the Nordgen accessions, Four were of Latvian origin and therefore excluded.
-GATK: 
-The reads were aligned to the Williams 82 2a reference genome ( (Schmutz et al.
+\textit{GATK:} The reads were aligned to the Williams 82 2a reference genome ( (Schmutz et al.
 2010).) adapted from GATK Best Practices, keeping MQ30 and biallelic sites and applied a MAF filter of 0.01 in order to remove rare variants from the data, reducing the number of SNPs from 17,648,123 to 10,000,122.
 For a detailed account of programs and commands used see supplementary material: \nameref{sec:supplementary:material} 
-From when i received it:
-A further three accessions were removed due to missing metadata and a single accessions with missing data > 5\% using BCFtools \cite{bcftools}, resulting in 153 nordgen accessions and 409 CCA accessions and a SNP count of 8,533,444.
+\textit{From when i received it:} A further three accessions were removed due to missing metadata and a single accessions with missing data > 5\% using BCFtools \cite{bcftools}, resulting in 153 nordgen accessions and 409 CCA accessions and a SNP count of 8,533,444.
 
-SNP data
-The snp data is a high-density genotyping array for soybean with 47,337 single nucleotide polymorphisms (SNPs) was develop from soybean (Glycine max L.
+\textit{SNP data:} The SNP data is a high-density genotyping array for soybean with 47,337 single nucleotide polymorphisms (SNPs) was develop from soybean (Glycine max L.
 Merr.).
@@
 Method
-pca
-Principal component analysis were applied using the package SnpRelate in R (A Parallel Computing Toolset for Relatedness and Principal Component Analysis of SNP Data).
+Principal Component Analysis (PCA) on SNP genotype data were applied using the package SnpRelate in R (A Parallel Computing Toolset for Relatedness and Principal Component Analysis of SNP Data).
+The functions in SNPRelate for PCA used (snpgdsPCA) calculate genetic covariance matrix from genotypes, computing the correlation coefficients between sample loadings and genotypes for each SNP.
+SNP eigenvectors were calculated.
+For the PCA the whole genome sequence was used for some of the PCA or the  intersect of the whole genome sequences the Core Collection accessions, the Swedish Nordgen accessions and the 50kSNP data available of the ten accessions identified as possible genetic Founders to the Swedish breeding program is used as indicated in the figure \ref{fig:pca}
 
-IBS Matrix: The package SNPRelate in R was used to compute an identity-by-state (IBS) matrix that calculates the proportion that two randomly selected reads that contain a certain SNP locus are the same or different between two individuals.
+The package SNPRelate in R was used to compute an identity-by-state (IBS) matrix that calculates the proportion that two randomly selected reads that contain a certain SNP locus are the same or different between two individuals.
 The resulting pairwise IBS matrix was used to generate a cluster dendogram using the function \textit{snpgdsHCluster: Hierarchical cluster analysis In SNPRelate }in R.
 
- snpgdsHCluster: Hierarchical cluster analysis
-In SNPRelate: Parallel Computing Toolset for Relatedness and Principal Component Analysis of SNP Data
-
-realisation that 2 of the swedish were likely not a part of the swedish breeding program and excluded 
-see Supplimentary material for accession data. 
-Fst of the swedish vs the cca
+The population differentiation measured in the populations the swedish accessions as a subpopulation and the core collection as a representation of the total population
+Fst is comparing the expected heterozygosity in the subpopulation (HS) to that expected in the total population HT :
+\(FST =1− (HS/HT)\)
 
@@
 H1: 
-To better understand the population relationships of the accessions received from the Nordic gene bank and how they are related to each other, and the rest of the soybean germplasm, they are compared to a core collection of 409 soybean from the USDA genebank selected from the soybean (Glycine max) germplasm for Central European breeding.
-The Core collection used here is a subset of the from a large soybean germplasm collection, the  USDA genebank accessions, that are likely preadapted to cultivation in Central Europe, while simultaneously conserving a high level of genetic diversity (CCA paper).
-Additionally accessions were grouped into categories based on known breeding history, So the suspected Founders of the Swedish breeding program are grouped together.
-Here an intersect of the whole genome sequences the Core Collection accessions, the Swedish Nordgen accessions and the 50kSNP data available of the ten accessions identified as possible genetic Founders to the Swedish breeding program is used.
+To better understand the population relationships of the accessions received from the Nordic gene bank and how they are related to each other, and the rest of the soybean germplasm , they are compared to a core collection of 409 soybean from the USDA genebank selected from the soybean (Glycine max) germplasm for Central European breeding.
+The Core collection used here is a subset of the soybean germplasm collection, the  USDA genebank accessions, which are assessed to be  preadapted to cultivation in Central Europe, while simultaneously conserving a high level of genetic diversity \cite{haupt20}.
+Additionally accessions were grouped into categories based on the suspected breeding history.
+There are the ten suspected Founders of the Swedish breeding program and the Swedish accessions of the Nordgen genebak.
 
-To understand the population differentiation/ diversity, a Principal Component Analysis (PCA) and the genome-wide Identity By State (IBS) pairwise distance matrix were applied.
-The resulting PCA shows the Swedish accessions as a sub group seen in green in (Figure \ref{fig:pca}) .
-A closer look at the Nordgen Swedish accessions reviled two accessions that were not part of the particular breeding program as seen in (Figure X showing the SBPA PCA with circles around the two accessions not from the breeding program)  and also thereafter confirmed by the genebank passport data.
-(Or do i simply remove these from the analysis and only mention that they are excluded up in the methods when describing the data)
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{plot_PCA_origin.pdf}
-\caption{This PCA shows the CCA in black, the SBPA in green and the thought to be founders in red.}%
-\label{fig:pca}
-\end{figure}
-
- and the IBS (Figure  \ref{fig:dendo}) shows . 
+A Principal Component Analysis (PCA) and the genome-wide Identity By State (IBS) pairwise distance matrix were applied to visualise the overall variance of the data.
+The resulting PCA (Figure \ref{fig:pca}) shows clustering of the Swedish accessions as a sub group based on similarity .
+A closer look at the Nordgen Swedish accessions reviled two accessions that were not part of the particular breeding program as seen in (Figure \ref{fig:pca3}  showing the PCA indicating the two accessions not from the breeding program in the cluster with the CCA in orange and black)  and also thereafter confirmed by the genebank passport data that they originated from other swedish soybean breeding programs (mention more here when certian.
+Altonagård and Ugra soja/ saja, malmåhus) .
+This figure also shows the accessions of the Nordgen and core collection that are swedish of origin and with a name consisting of letters (excluding numbers or non-alphanumeric charachters of theaccessions names) if included in the data.
+The IBS (Figure  \ref{fig:dendo}) clearly shows the swedish accessions of the breeding program of interest clustering together. The  
  IBS numbers in a table showing ibs within sbpa and out?
 
-The genetic differentiation measured by FST
+The population differentiation measured by Fst
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{rainbow.pdf}
-\caption{Hierarchical cluster dendrogram based on pairwise identity-by-state (IBS) values from WGS data for all samples. }%
-\label{fig:dendo}
-\end{figure}
+postulate: And population subdivision analysis will quantify the SPBA will be a distinct subpopulation with a moderate to distinct allele frequency difference.
 
 \subsection{Genetic diversity} 
-
-MAF?
-LD
-pi 
-theta
+H2
 
 \subsection{Genome wide selection signatures} 
-gws.
+H3
 
@@
 
-\section{Additional guidelines}
-
-\subsection{Numbers} In the text, write out numbers nine or less except as part of a date, a fraction or decimal, a percentage, or a unit of measurement.
-Use Arabic numbers for those larger than nine, except as the first word of a sentence; however, try to avoid starting a sentence with such a number.
-
-\subsection{Units} Use abbreviations of the customary units of measurement only when they are preceded by a number: "3 min" but "several minutes".
-Write "percent" as one word, except when used with a number: "several percent" but "75\%." To indicate temperature in centigrade, use ° (for example, 37°); include a letter after the degree symbol only when some other scale is intended (for example, 45°K).
-
-\subsection{Nomenclature and italicization} Italicize names of organisms even when  when the species is not indicated.
-Italicize the first three letters of the names of restriction enzyme cleavage sites, as in HindIII.
-Write the names of strains in roman except when incorporating specific genotypic designations.
-Italicize genotype names and symbols, including all components of alleles, but not when the name of a gene is the same as the name of an enzyme.
-Do not use "+" to indicate wild type.
-Carefully distinguish between genotype (italicized) and phenotype (not italicized) in both the writing and the symbolism.
-
-\subsection{Cross references}
-Use the \verb|\nameref| command with the \verb|\label| command to insert cross-references to section headings.
-For example, a \verb|\label| has been defined in the section \nameref{sec:materials:methods}.
-
-\section{In-text citations}
-
-Add citations using the \verb|\citep{}| command, for example \citep{neher2013genealogies} or for multiple citations, \citep{neher2013genealogies, rodelsperger2014characterization,Falush16}
-
-\section{Examples of article components}
-\label{sec:examples}
-
-The sections below show examples of different header levels, which you can use in the primary sections of the manuscript (Results, Discussion, etc.) to organize your content.
-
-\section{First level section header}
-
-Use this level to group two or more closely related headings in a long article.
-
-\subsection{Second level section header}
-
-Second level section text.
-
-\subsubsection{Third level section header:}
-
-Third level section text. These headings may be numbered, but only when the numbers must be cited in the text.
-
 \section{Figures and tables}
@@
 
-\subsection{PCA figure}
+\subsection{PCA1 figure}
 
@@
 \includegraphics[width=\linewidth]{plot_PCA_origin.pdf}
-\caption{This PCA shows the CCA in black, the SBPA in green and the thought to be founders in red.}%
+\caption{Principal component analysis for soybean samples from defined groups, Core collection Accessions (CCA) in black, the Swedish Breeding Program Accessions (SBPA) in green and the Founders of the breeding progran (F) in red.}%
 \label{fig:pca}
 \end{figure}
+\subsection{PCA2 figure}
 
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics[width=\linewidth]{example-figure}
-%%\caption{Example movie (the figure file above is used as a placeholder for this example).
-\textit{GENETICS} supports video and movie files that can be linked from any portion of the article - including the abstract.
-Acceptable formats include .asf, avi, .wav, and all types of Windows Media files.
-%%}%
-%%\label{video:spectrum}
-%\end{figure}
+Figure \ref{fig:pca2} .
 
-Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{plot_PCA_mg1.pdf}
+\caption{Principal component analysis for soybean samples from the whole genome sequence data, including the 155 Nordgen accessions and the 409 Core collection accessions.
+Indicated is the Maturity groups assigned to the accessions.}
+\label{fig:pca2}
+\end{figure}
 
+\subsection{PCA3 figure}
+
+Figure \ref{fig:pca3} .
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{plot_PCA_flag.pdf}
+\caption{PCA MG status}%
+\label{fig:pca3}
+\end{figure}
 \subsection{Dendogram figure}
@@
 \end{figure}
-
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics[width=\linewidth]{example-figure}
-%%\caption{Example movie (the figure file above is used as a placeholder for this example).
-\textit{GENETICS} supports video and movie files that can be linked from any portion of the article - including the abstract.
-Acceptable formats include .asf, avi, .wav, and all types of Windows Media files.
-%%}%
-%%\label{video:spectrum}
-%\end{figure}
-
-Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
 
@@
 %\end{figure}
+
+Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
 
@@
 Supplementary Material Methods (S Methods)
+\href{PCA.html}
 
@@
 
+\section{In-text citations}
+
+Add citations using the \verb|\citep{}| command, for example \citep{neher2013genealogies} or for multiple citations, \citep{neher2013genealogies, rodelsperger2014characterization,Falush16}
+
 \bibliography{bibliography}

### 2023-07-31 manuscript_soy.tex
Backup

@@
 \usepackage{epstopdf}
-
-\articletype{gs} % article type
-% {inv} Investigation
-% {gs} Genomic Selection
-% {goi} Genetics of Immunity
-% {gos} Genetics of Sex
-% {mp} Multiparental Populations
-
-\runningtitle{Soy manuscript} % For use in the footer
+\usepackage[T1]{fontenc}
+\usepackage[latin1]{inputenc}
+\articletype{inv}
+\runningtitle{Soy scientific manuscript} % For use in the footer
 \runningauthor{Connelly \textit{et al.}}
 
-\title{Working title SoyAdapt}
+\title{SoyAdapt}
 
-\author[1,$\dagger$,$\ast$]{Josephine Estelle Ananda Connelly}
-\author[2]{Guillaume Paul Ramstein}
-\author[3]{Wolf L Eiserhardt}
-\author[4]{Torben Asp}
+\author[1,$\ast$]{Josephine Estelle Ananda Connelly}
+\author[2,$\dagger$]{Guillaume Paul Ramstein}
+\author[3,$\dagger$]{Wolf L Eiserhardt}
+\author[4,$\dagger$]{Torben Asp}
 
@@
 
-\correspondingauthoraffiliation[$\ast$]{Corresponding author: AU, JosephineConnelly@qgg.uni.dk.}
+\affil[$\dagger$]{These authors share senior authorship}
+\correspondingauthoraffiliation[$\ast$]{Corresponding author: AU, josephineconnelly@qgg.au.dk.}
 
 \begin{abstract}
-Soybean (Glycine max L. Merrill) is the worlds leading  oilseed crop used as a  primary source of vegetable oil for human consumption and in proteinmeal for animal feed.
+Soybean (Glycine max L. Merrill) is the worlds leading  oilseed crop used as a  primary source of vegetable oil for human consumption and in protein meal for animal feed.
+
+Search engine optimization (SEO)
+
+There are a few simple ways to maximize your article's discoverability and search results.
+
+    Include a few of your article's keywords in the title of the article
+
+    Do not use long article titles
+
+    Pick 5-8 keywords using a mix of generic and more specific terms on the article subject(s)
+
+    Use the maximum amount of keywords in the first two sentences of the abstract
+
+    Use some of the keywords in level 1 headings
+
+    Title
+
+The title should be concise, omitting terms that are implicit and, where possible, be a statement of the main result or conclusion presented in the manuscript.
+Abbreviations should be avoided within the title.
+
+Witty or creative titles are welcome, but only if relevant and within measure.
+Consider if a title meant to be thought-provoking might be misinterpreted as offensive or alarming.
+In extreme cases, the editorial office may veto a title and propose an alternative.
+
+Authors should avoid:
+
+    titles that are a mere question without giving the answer
+
+    unambitious titles, for example starting with 'Towards,' 'A description of,' 'A characterization of' or 'Preliminary study on'
+
+    vague titles, for example starting with 'Role of', 'Link between', or 'Effect of' that do not specify the role, link, or effect
+
+    including terms that are out of place, for example, the taxonomic affiliation apart from species name.
+
+For Corrigenda, General Commentaries, and Editorials, the title of your manuscript should have the following format.
+
+    'Corrigendum: [Title of original article]'
+
+    General Commentaries:
+    'Commentary: [Title of original article]'
+    'Response: Commentary: [Title of original article]'
+
+    'Editorial: [Title of Research Topic]'
 \end{abstract}
@@
 \thispagestyle{firststyle}
-%\slugnote
-%\firstpagefootnote
-\vspace{-13pt}% Only used for adjusting extra space in the left column of the first page
-
-\section{Abstract}
-
-Soybean, belonging to the Fabaceae family and scientifically known as \textit{Glycine max} (L.
-Merrill), is globally recognised as the primary legume crop due to its high protein and oil content.
-However, its genetic diversity has decreased over time due to domestication and intensive artificial selection, driving the need to look elsewhere to increase germplasm utilisation for abiotic adaptation \citep{hyten06, gizlice96}.
-Relevant accessions with the potential for abiotic adaptation to cold tolerance have been identified [@haupt20].
-Additionally 160 Swedish accessions originating from a Swedish breed for the adaptation to the cool climate of north western Europe and stored in the NordGen gene bank have been unearthed [@holmberg1973].
-We will interpret the plant genetic resources available, so as to utilise the genetic potential.
-To facilitate targeted selection of promising accessions from soybeans vast germplasm collections,  and unveiling the genetic data stored in the NordGen seed banks.
-findings are: ...
+\vspace{-13pt}
 
@@
 
-\textit{soybean needs to be grown in the north}:  This argument but then for spread of growth area!
-"Modern U.S.
-soybean breeding has led to a yield increase of 29 kg ha/−1 yr/−1 (Rincker et al., 2014)"
- "Better understanding the genomic basis behind this improvement may provide indicators for further soybean improvement and adaptation.
-"  from: Sequencing the USDA core soybean collection reveals gene loss during domestication and breeding"
+\textit{soybean needs to be grown in the north}:
+Modern soybean breeding has led to yield increases of $229 \, \text{kg} \, ha^{-1} \, \text{yr}^{-1}$ (Rincker et al., 2014), gaining a deeper understanding of the genomic basis of this progress could offer valuable insights for advancing and adapting soybean cultivation even further.
 
-\textit{the swedish soybeans (goals and scope)}
+Expanding the production of soy to the northern climates by leveraging the advancements in soybean breeding and focusing on the genes relevant to adaption, can diversify the global supply chain, reduce dependency, and potentially alleviate the environmental pressures associated with soy production in high biodiversity, sensitive areas/ecologically vulnerable regions.
+With the global demand for soy and the need for sustainable agricultural practices, investing in breeding programs to develop soybean varieties suitable for northern climates can present a promising solution.
+
+\textit{the Swedish soybeans (goals and scope)}
 
@@
 \textit{In this study we}
-The goal is to help precipitate the utilisation of these soybean genetic resources.
+The goal is to help precipitate the utilisation of these soybean genetic resources. to increase germplasm utilization for through understanding what we have in the genebanks.
 
@@
 
-what is the data: 160 soybean accessions were obtained from the Nordgen genebank, 155 of these accessions are of Swedish origin.
-These accessions are thought to originate from a Swedish breeding program running from the 1840s to the 1970s which used material consisting of a mixture of elite cultivars of the time and Japanese landraces, breeding for the adaptation to the cool climate of north western Europe \cite{holmberg1973}.
+\subsection{Swedish Breeding Program Collection}
+155 soybean accessions were obtained from the Nordgen genebank of Swedish origin.
+These accessions originate from a Swedish breeding program (SBP) running from the 1840s to the 1970s which used material consisting of a mixture of elite cultivars of the time and Japanese landraces, breeding for the adaptation to the cool climate of northwestern Europe \cite{holmberg1973}.
 
-The Core collection used here is a subset of the from a large soybean germplasm collection, the  USDA genebank accessions consisting of 415 of the USDA soy germplasm accessions.
-This Core collection has been selected with a focus on adaptation to high-latitude cold regions using environmental data from phenotypic trials in Germany, and comparing Donor opulation of Environments (DPE) in Asia and the Target Population of Environments (TPE) in Central Europe.
-From the 3663 accessions two diverse core collections of 183 and 366 accessions were created.
-These 514 diversity panels are used here due to that they are likely preadapted to cultivation in Central Europe, while simultaneously conserving a high level of genetic diversity .
+\subsection{The Core Collection}
+The Core Collection (CCA) consists of two samplings of the core collection from haupt20. A 5\% and 10\% collection with overlapping accessions in total consisting of 409 accessions after filtering. 
+The Core collection accessions (CCA) used here is a subset of the large soybean germplasm collection, the  USDA genebank accessions consisting of 415 of the USDA soy germplasm accessions.
+This Core collection has been selected with a focus on adaptation to high-latitude cold regions using environmental data from phenotypic trials in Germany, and comparing Donor opulation of Environments (DPE) in Asia and the Target Population of Environments (TPE) in Central Europe by \cite{haupt20}.
+From the more than 17.000 accessions, two diverse core collections of 183 and 366 accessions were created.
+These 514 diversity panels are used here due to that they are likely preadapted to cultivation in Central Europe, while simultaneously conserving a high level of genetic diversity.
 
-USDA germplasm of soybean consists of ... soybase.org.  Founders 10.
+\subsection{The Founders Collection}
+The USDA Soybean Germplasm Collection maintains accessions of >17000 cultivated soybeans and is part of the U.S.
+National Plant Germplasm System (NPGS).
+These accessions have been previously genotyped with the SoySNP50K iSelect BeadChip  \cite{song13,song15} has been used for genotyping the USDA Soybean Germplasm Collection and is available for use from SoyBase (https://soybase.org/) in the version mapped to the soybean genome assembly Glyma.Wm82.a2  \cite{schmutz10}.
+In the pedigree information from Nordgen SBPA, are 10 accessions present in the USDA genebank and thus are part of the SoySNP50K genotypic data.
+These specific accessions were included as they are considered Founders of the SBP.
 
-wgs
-\textit{What was done:} The accessions from the Nordgen genebank and the Core collection accessions were whole genome sequenced.
+\textit{DNA extraction and sequencing information?:}
+
+\subsection{Genotyping and filtering of whole genome sequence SNP}
+The accessions from the Nordgen genebank and the Core collection accessions were whole genome sequenced.
 Four of the Core collection seeds didn't germinate in time for the sequencing, resulting in accessions sequenced 574.
-Of the Nordgen accessions, Four were of Latvian origin and therefore excluded.
-\textit{GATK:} The reads were aligned to the Williams 82 2a reference genome ( (Schmutz et al.
-2010).) adapted from GATK Best Practices, keeping MQ30 and biallelic sites and applied a MAF filter of 0.01 in order to remove rare variants from the data, reducing the number of SNPs from 17,648,123 to 10,000,122.
-For a detailed account of programs and commands used see supplementary material: \nameref{sec:supplementary:material} 
-\textit{From when i received it:} A further three accessions were removed due to missing metadata and a single accessions with missing data > 5\% using BCFtools \cite{bcftools}, resulting in 153 nordgen accessions and 409 CCA accessions and a SNP count of 8,533,444.
+Of the Nordgen accessions, Four were of Latvian origin and therefore excluded in further analysis.
+The reads were aligned to the Williams 82 2a reference genome \cite{schmutz10} using an adapted GATK Best Practices \cite{mckenna10}, keeping MQ30 and biallelic sites and using VCFTools \cite{danecek11} applying a minor allele frequency (MAF) filter of 0.01 in order to remove rare variants from the data, reducing the number of SNPs markers were initially called (raw dataset) from 17,648,123 to 10,000,122.
+For a detailed account of accessions, programs and commands used see supplementary material: \nameref{sec:supplementary:material} 
+A further three accessions were removed due to missing phenotype data and a single accession that had missing data > 5$\%$ using BCFtools \cite{danecek21}, resulting in 153 Swedish accessions from the NordGen genebank and 409 of the CCA accessions.
+Five per cent missing data at SNPs was also removed resulting in an SNP count of 8,533,444.
+This data set was used to compare the Allele Frequency of the two populations.
+For all other analyses, MAF 5\%  was removed using Vcftools (--maf 0.05) on the dataset as a whole, except for the case of LD calculation MAF5\% was removed separately for SBPA and CCA.
 
-\textit{SNP data:} The SNP data is a high-density genotyping array for soybean with 47,337 single nucleotide polymorphisms (SNPs) was develop from soybean (Glycine max L.
-Merr.).
-The SoySNP50K iSelect BeadChip has been used to genotype the USDA Soybean Germplasm Collection and is available for use from soybase
- An intersect was made of the whole genome sequenced
-35486
+\subsection{SNP intersect }
+An intersection was made of the whole genome sequenced data (153 SBPA and 409 CCA) and the SoySNP50K array data of the 10 Founders.
+This was done using BCFtools with the command "isec".
+The intersecting data was 35.486 SNPs.
 
-For a detailed account of programs and commands used see supplementary material: \nameref{sec:supplementary:material}
+For a detailed account of programs and commands, see supplementary material, \nameref{sec:supplementary:material} (S3: Methods 1) data preparation and filtering and \nameref{sec:data:availability} for data.
 
-H1
-Q 
+\subsection{Population Structure Analysis}
 To infer population stratification of the Swedish accessions from the NordGen gene bank and the Core collection and of the 12 accessions assumed to be founders of the Swedish breeding program.
-
-Method
-Principal Component Analysis (PCA) on SNP genotype data were applied using the package SnpRelate in R (A Parallel Computing Toolset for Relatedness and Principal Component Analysis of SNP Data).
-The functions in SNPRelate for PCA used (snpgdsPCA) calculate genetic covariance matrix from genotypes, computing the correlation coefficients between sample loadings and genotypes for each SNP.
-SNP eigenvectors were calculated.
-For the PCA the whole genome sequence was used for some of the PCA or the  intersect of the whole genome sequences the Core Collection accessions, the Swedish Nordgen accessions and the 50kSNP data available of the ten accessions identified as possible genetic Founders to the Swedish breeding program is used as indicated in the figure \ref{fig:pca}
+Principal component analysis (PCA) on SNP genotype data was applied using the package SnpRelate in R (A Parallel Computing Toolset for Relatedness and Principal Component Analysis of SNP Data).
+The functions in SNPRelate for PCA used (snpgdsPCA) calculate the genetic covariance matrix from genotypes, computing the correlation coefficients between sample loadings and genotypes for each SNP and SNP eigenvectors were calculated.
+The WGS data was used for a number of the PCA, and where the possible genetic Founders to the Swedish breeding program the intersect data is used and indicated in figures; \ref{fig:pca}.
 
 The package SNPRelate in R was used to compute an identity-by-state (IBS) matrix that calculates the proportion that two randomly selected reads that contain a certain SNP locus are the same or different between two individuals.
-The resulting pairwise IBS matrix was used to generate a cluster dendogram using the function \textit{snpgdsHCluster: Hierarchical cluster analysis In SNPRelate }in R.
+The resulting pairwise IBS matrix was used to generate a cluster dendrogram using the function \textit{snpgdsHCluster: Hierarchical cluster analysis In SNPRelate }in R.
 
-The population differentiation measured in the populations the swedish accessions as a subpopulation and the core collection as a representation of the total population
-Fst is comparing the expected heterozygosity in the subpopulation (HS) to that expected in the total population HT :
-\(FST =1− (HS/HT)\)
+Relatedness measures F-statistics (also known as fixation indices) for given populations were calculated using the function \textit{snpgdsFst: F-statistics (fixation indices) in SNPRelate }in R.
+The whole genome data was used to calculate the pairwise population differentiation measured in the populations of the Swedish accessions as a subpopulation with the Core Collection as a representation of the total population.
+Fst is comparing the expected heterozygosity in the subpopulation (HS) to that expected in the total population (HT) :
 
-H2 
-Q diversity
+$F_{\text{ST}} = 1 - \left(\frac{H_{\text{S}}}{H_{\text{T}}}\right)$
 
-Method
-LD
-pi and
+Later in the analysis when looking for signatures of selection Fst is calculated according to the methods of \cite{Weir84} using the Vcftools \cite{danecek11} function --weir-fst-pop on the SNP variants with (no window size) a window size of 1 Mb with a step size of 100 kb.
 
-H3
-Q
+\subsection{Allele frequency, LD and nucleotide diversity}
+The idea is to look into the effects of the bottleneck caused by the SBP, the increased drift and will be detected as a reduction of Ne.
+Detecting a decrease in genetic diversity as a measure of allele frequency, nucleotide diversity and LD increase.
+Allele Frequency (AF) was calculated using Vcftools --freq and plotted in R.
+For the LD analysis vcftools --maf 0.05 for each population was removed separately.
+The function \textit{snpgdsLDmat: Linkage Disequilibrium (LD) analysis, in SNPRelate } in R.
+A matrix of "r" for R coefficient (by EM algorithm assuming HWE) is calculated for pairwise snps separately for each chromsome.
+The squared correlation coefficient r2 is a popular measure of LD an chosen due to the advantage of this calculation being less sensitive to marginal allele frequencies over  D or D' \cite{sved18}.
 
-Method
+\(r^2 = \frac{D^2}{p_Ap_B(1 - p_A)(1 - p_B)}\) where pA and pB are the allele frequencies and D is the coefficient of LD:
+
+$D=pAB−pApB$
+
+The $r^2$  values were binned and plotted using ggplot2 \cite{wickham16} in R (ref).
+Nucleotide diversity ($\pi$), measuring the degree of polymorphism in the populations was calculated with the wgs data using vcftools for 100 kb windows --window-pi 10000 \cite{nei79}.
+*This is appropriate @ 25kb as we just discussed earlier today with the fst.
+
+\subsection{H3}
+Q and Method
 
@@
 \section{Results and Discussion}
-\subsection{population structure and genetic differentiation}
 
-H1: 
-To better understand the population relationships of the accessions received from the Nordic gene bank and how they are related to each other, and the rest of the soybean germplasm , they are compared to a core collection of 409 soybean from the USDA genebank selected from the soybean (Glycine max) germplasm for Central European breeding.
-The Core collection used here is a subset of the soybean germplasm collection, the  USDA genebank accessions, which are assessed to be  preadapted to cultivation in Central Europe, while simultaneously conserving a high level of genetic diversity \cite{haupt20}.
-Additionally accessions were grouped into categories based on the suspected breeding history.
-There are the ten suspected Founders of the Swedish breeding program and the Swedish accessions of the Nordgen genebak.
+\subsection{H1 The Swedish soybean accessions were part of the SBP}
+The accessions from the NordGen gene bank (Swedish breeding program accessions, SBPA) are closely related and therefore we can conclude that they were part of the same breeding program, the Swedish breeding program (SBP), of the 1940s-1980s of Algot Holmberg and Sons from Norrköping, producing the known cultivars, Fiskeby III, Fiskeby V, Träff $\&$ Bråvalla, which are included in the NordGen gene bank accessions and in this analysis.
+Two accessions are identified as accessions that were not part of the Swedish breeding program of interest but have been part of other Swedish or Nordic soybean breeding trajectories and later obtained from the Russian seed bank.
 
-A Principal Component Analysis (PCA) and the genome-wide Identity By State (IBS) pairwise distance matrix were applied to visualise the overall variance of the data.
-The resulting PCA (Figure \ref{fig:pca}) shows clustering of the Swedish accessions as a sub group based on similarity .
-A closer look at the Nordgen Swedish accessions reviled two accessions that were not part of the particular breeding program as seen in (Figure \ref{fig:pca3}  showing the PCA indicating the two accessions not from the breeding program in the cluster with the CCA in orange and black)  and also thereafter confirmed by the genebank passport data that they originated from other swedish soybean breeding programs (mention more here when certian.
-Altonagård and Ugra soja/ saja, malmåhus) .
-This figure also shows the accessions of the Nordgen and core collection that are swedish of origin and with a name consisting of letters (excluding numbers or non-alphanumeric charachters of theaccessions names) if included in the data.
-The IBS (Figure  \ref{fig:dendo}) clearly shows the swedish accessions of the breeding program of interest clustering together. The  
- IBS numbers in a table showing ibs within sbpa and out?
+\subsubsection{Relatedness: The SBPA are genetically similar and therefore closely related}
+There are 155 soybean accessions sequenced from the Nordgen genebank.
+153 of the accessions are similar to a degree that we can conclude they are closely related to a degree of XXX.
+numbers from the ibs?
+Two of the Swedish accessions (Altonagård \& Ugra Soja) show up clearly distinct from the cluster of SBPA and are therefore no longer considered or included as accessions of the investigated (SBP).  
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{plot_heatmap.pdf}
+\caption{This heatmap includes Altonagård and Ugra soja which are seen as the bottom two blue acessions. They are dissimilar to eachother and dissimmilar to the rest of the accessions.}
+\label{fig:heat}
+\end{figure}
 
-The population differentiation measured by Fst
+Characterise genetic relatedness. 
+PCA the first 10 pcs capture 22.82\%  of the variation.
+the ibs shows identity towards 1 to be at: numbers?
 
-postulate: And population subdivision analysis will quantify the SPBA will be a distinct subpopulation with a moderate to distinct allele frequency difference.
+\subsubsection{Distinctness: The SBPA are a genetically cohesive group}
+The SBPA are a genetically cohesive group of accessions.
+In comparison to the CCA, they are similar to each other (matrix numbers?) as seen in the figure the SBPA cluster within the dendrogram of the CCA and the SBPA.
+It is also clear that some of the CCA are closely related to the SBPA, which is expected as a domesticated crop and the history known of other pedigrees and breeding programs exchanging germplasm.
+In the CCA, 94 of the accessions are of European origin and additionally, seven of those  originate from Sweden, some with accession names identical or similar to that of the SBPA.
+\begin{figure}
+\centering
+\includegraphics[width=\linewidth]{rainbow.pdf}
 
-\subsection{Genetic diversity} 
-H2
+\caption{Hierarchical cluster dendrogram based on pairwise identity-by-state (IBS) values from the 50K SNP data for all samples.
+Including the suspected Founders.
+The dark blue accessions grouped together are a genetically cohesive group.
+Based on An IBS analysis is seen that the SBPA are genetically distinct from the other accessions, clustering together.
+Several Founders are seen clustering hierarchically outside the SBPA.
+The Swedish accessions Altonagård and Ugra Soja cluster outside }%
+\label{fig:dendo}
+\end{figure}
 
-\subsection{Genome wide selection signatures} 
-H3
+\subsubsection{Divergence: There is a (significant) divergence within the subpopulation SBPA of the soybean when comparing other groups of soybean}
 
-\section{Discussion}
+Genetic differentiation, e.g., Fst was calculated using SBPA as a subpopulation of the CCA.
+This is argumentatively valid since the CCA has a variability comparable to the soybean germplasm as a whole (ref).
+Fst mean at 0.178
+The CCA has furthermore been grouped into subpopulations according to Maturation score status to compare the divergence of the SBPA. 
+Figure: Fst
+The high Fst of the SBPA shows that there is a considerable degree of differentiation in comparison to the other groups defined by maturity group shows
 
-The Swedish soybeans are are mainly a snapshot of a breeding program stopped around 1978 and all lines or recent crosses from this program were then set into the
+A relatively high Fst mean between (0.405 \& 0.215) for the SBPA in comparison to the CCA groups shows that there is a considerable degree of differentiation in relation to the other populations as defined by the maturity group.
+The CCA in relation to each other has a relatively low Fst, between (0.075 \& 0.007) showing that differentiation is not high when grouping the CCA by Maturity.
+
+\textbf{Founders}
+p4. summary of what we can see of the founders. 
+A list of known and suspected founders has been included in the PCA.
+It is clear that (some specific) founders are closely related to the SBPA and were clearly founders of the SBP whereas others are not distinguishable as Founders.
+It is expected that the founder accessions could be more related to accessions in the CCA than the SBPA.
+A more detailed picture of decent can be seen in the supplementary material where the pedigree data of the SBPA are collected.
+\nameref{sec:supplementary:material}.
+
+Figure: PCA with founders in. 
+The founders could also be indicated earlier in the dendrogram clustering together with the SBPA and the SBPA cluster nested within the founders.
+
+\subsection{H2 Genetic diversity}
+
+\textbf{The SBP caused a bottleneck as seen in a decrease in genetic diversity }
+
+There is a reduction of Ne detected in a high level of linkage of the SBPA, the lower nucleotide diversity ($\pi$ ) There is also an accumulation of rare alleles and a lot of alleles fixed in the population.
+These observations suggest that the bottleneck event and the ensuing intensification of genetic drift had a noticeable impact on the genetic composition of the SBP population.
+
+Earlier studies have shown a halving in nucleotide diversity when comparing wild soybean to their domesticated progenitors and a further halving in nucleotide diversity in breeding regimes in recent years comparing landraces to elite cultivars (ref).
+The SBPA have a mean nucleotide diversity of ($\pi$ = 0.000999), 0.09\% or a difference of  9/10.000 bases between two sequences.
+Whereas the CCA has almost double the mean nucleotide diversity of  ($\pi$ = 0.001911)  0.19\% or 19/10.000 bases differing.
+The CCA has two times higher nucleotide diversity than the SBPA.
+This is consistent with a bottleneck during the selection of founders and subsequent selection by the breeding program.
+
+The nonrandom association of alleles at different sites due to recombination (LD) in soybean shows like in the CCA
+In this comparison of CCA and SBPA, we see a high LD at both low and high ranges of the SBPA.
+The LD slope of the SBPA shows a higher degree of decay over shorter physical distances, indicating in relation to the CCA, a contrasting population size.
+
+Longer range association estimates a small Ne, because of the bottleneck. 
+The long-range LD will be due to a high level of fixation caused by drift.
+
+figure text:  The LD plot shows the relationship between genetic markers in the two populations at different sites due to recombination.
+The slope represents the rate at which LD decays with physical distance.
+The top cyan line is SBPA and the red line is the CCA.
+
+A difference in slope can also indicate different selective pressures or adaptive processes, such as the steep and higher slope of the SBPA may suggest positive selection or recent selective sweeps.
+
+AF 
+The fixation of alleles due to drift is also seen in the allele frequency of the SBPA. The high number of fixated alleles seen as 
+in a high number of allele frequencies in the extremities of the frequency spectrum is a pattern of fixation due to drift.
+Increased low freq can also be a signal of negative directional selection or a selective sweep, and an increase in high-frequency variants can be a signal of positive directional selection \cite{}.
+
+\textbf{Allele Frequency shows a high level of drift}
+The high number of alleles at high frequency shows that the SBPA have a higher number of monomorphic locus than the CCA.
+This high number of fixed alleles arises through natural selection, genetic bottlenecks, or genetic isolation in the population.
+The SBPA also has a large number of rare alleles, a high number of genetic variants in low frequencies.
+Rare alleles arise through mutation, genetic drift, or migration from other populations.
+
+The presence of numerous rare alleles, seen in combination with the low pi and high decay, shows the genetic drift in the population. 
+or
+In conjuncture, the high number of rare alleles and high number of fixed alleles together reveal the bottleneck as also seen in the following LD decay analysis.
+
+We can deduce this since we know we are not looking at a diverse genetic pool, which could otherwise be a signal of a recent population expansion or balancing selection.
+Since we conclude that the accumulation of rare alleles is due to drift, we can assume that there is an accumulation of deleterious alleles in a population: The deleterious alleles can arise through spontaneous mutations and random fluctuations in allele frequencies, and the small size of the population leads to an increased frequency of harmful variants.
+
+Although the findings above raise the possibility of potential selection processes occurring in the SBP, the high effect of genetic drift could mask any such signatures of selection making them indistinguishable because the drift enhances random fluctuations in  frequency making it difficult to detect or for the allele to reach fixation.
+
+The next step is to predict if there are areas of the genome where the nucleotide diversity is lower than expected.
+
+\subsection{H3 Genome-wide selection signatures}
+
+Selection happened during the Swedish breeding program.
+Selection during the Swedish breeding program occurred and can be detected as the accessions are insensitive to photoperiod, and selection occurred signatures are detected at other genomic regions related to seed coat and hilum colour, cold tolerance, and flowering.
+
+Selection at genes of interest.
+ Selection can have a similar effect on LD in the genome as a bottleneck can, but it is a localised phenomenon in and around the targeted loci.
+We know the breeding program goals and have selected areas to target looking for selection signatures by using areas of the genome that are connected or hypothesised to be relevant to the expression involved to named traits.
+xpclr results.
+
+haplotype groups?
+
+The Swedish soybeans are mainly a snapshot of a breeding program stopped around 1978 and all lines or recent crosses from this program were then set into the
 
 \section{Figures and tables}
-
-Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
 
@@
 \centering
-\includegraphics[width=\linewidth]{plot_PCA_origin.pdf}
-\caption{Principal component analysis for soybean samples from defined groups, Core collection Accessions (CCA) in black, the Swedish Breeding Program Accessions (SBPA) in green and the Founders of the breeding progran (F) in red.}%
+\includegraphics[width=\linewidth]{plot_PCA_source.pdf}
+\caption{Biplot of the principal component analysis of soybean genotypes.
+The colours represent the defined groups; Core collection Accessions (CCA) in black, the Swedish Breeding Program Accessions (SBPA) in green and the Founders of the breeding program (F) in red.}%
 \label{fig:pca}
 \end{figure}
+
 \subsection{PCA2 figure}
@@
 \caption{Principal component analysis for soybean samples from the whole genome sequence data, including the 155 Nordgen accessions and the 409 Core collection accessions.
-Indicated is the Maturity groups assigned to the accessions.}
+Indicated are the Maturity groups assigned to the accessions.}
 \label{fig:pca2}
@@
 \includegraphics[width=\linewidth]{plot_PCA_flag.pdf}
-\caption{PCA MG status}%
+\caption{PCA flag}
 \label{fig:pca3}
 \end{figure}
+
+\subsection{PCA figure4}
+
+Figure \ref{fig:pca4} .
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{plot_PCA_region2.pdf}
+\caption{PCA regions}
+\label{fig:pca4}
+\end{figure}
+
 \subsection{Dendogram figure}
@@
 
-\subsection{Sample figure}
+\subsection{Heatmap figure}
 
-Figure \ref{fig:spectrum} shows an example figure.
+Figure \ref{fig:heat} .
 
@@
 \centering
-\includegraphics[width=\linewidth]{example-figure}
-\caption{Example figure from \url{10.1534/genetics.114.173807}.
-Please include your figures in the manuscript for the review process.
-You can upload figures to Overleaf via the Project menu.
-Images of photographs or paintings can be provided as raster images.
-Common examples of raster images are .tif/.tiff, .raw, .gif, and .bmp file types.
-The resolution of raster files is measured by the number of dots or pixels in a given area, referred to as “dpi” or “ppi.”
-\begin{itemize}
-\item  minimum resolution required for printed images or pictures: 350dpi
-\item  minimum resolution for printed line art: 600dpi (complex or finely drawn line art should be 1200dpi)
-\item minimum resolution for electronic images (i.e., for on-screen viewing): 72dpi
-\protect\end{itemize}
-Images of maps, charts, graphs, and diagrams are best rendered digitally as geometric forms called vector graphics.
-Common file types are .eps, .ai, and .pdf.
-Vector images use mathematical relationships between points and the lines connecting them to describe an image.
-These file types do not use pixels; therefore resolution does not apply to vector images.
-Label multiple figure parts with A, B, etc.
-in bolded.
-Legends should start with a brief title and should be a self-contained description of the content of the figure that provides enough detail to fully understand the data presented.
-All conventional symbols used to indicate figure data points are available for typesetting; unconventional symbols should not be used.
-Italicize all mathematical variables (both in the figure legend and figure) , genotypes, and additional symbols that are normally italicized.}%
-\label{fig:spectrum}
+\includegraphics[width=\linewidth]{plot_heatmap.pdf}
+\caption{PCA regions}
+\label{fig:heat}
 \end{figure}
 
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics[width=\linewidth]{example-figure}
-%%\caption{Example movie (the figure file above is used as a placeholder for this example).
-\textit{GENETICS} supports video and movie files that can be linked from any portion of the article - including the abstract.
-Acceptable formats include .asf, avi, .wav, and all types of Windows Media files.
-%%}%
-%%\label{video:spectrum}
-%\end{figure}
+\subsection{Table 1}
 
-Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
+Table  \ref{tab:fst} shows table 1.
 
-\subsection{Sample table}
+\section{other text}
 
-Table \ref{tab:shape-functions} shows an example table.
-Avoid shading, color type, line drawings, graphics, or other illustrations within tables.
-Use tables for data only; present drawings, graphics, and illustrations as separate figures.
-Histograms should not be used to present data that can be captured easily in text or small tables, as they take up much more space.
+abstract:
+Soybean, belonging to the Fabaceae family and scientifically known as \textit{Glycine max} (L.
+Merrill), is globally recognised as the primary legume crop due to its high protein and oil content.
+However, its genetic diversity has decreased over time due to domestication and intensive artificial selection, driving the need to look elsewhere to increase germplasm utilisation for abiotic adaptation \citep{hyten06, gizlice96}.
+Relevant accessions with the potential for abiotic adaptation to cold tolerance have been identified [@haupt20].
+Additionally 160 Swedish accessions, originating from a Swedish breed for the adaptation to the cool climate of north western Europe and stored in the NordGen gene bank have been unearthed [@holmberg1973].
+We will interpret the plant genetic resources available, so as to utilise the genetic potential.
+To facilitate targeted selection of promising accessions from soybeans vast germplasm collections,  and unveiling the genetic data stored in the NordGen seed banks.
+findings are: ...
 
-Tables numbers are given in Arabic numerals. Tables should not be numbered 1A, 1B, etc., but if necessary, interior parts of the table can be labeled A, B, etc. for easy reference in the text.
+\textbf{Founders}
+Founders identified through pedigrees and other genebank metadata have differing similarities to the SBPA.
 
-\begin{table*}[p]
+\textbf{Ugra soja}
+Ugra soja 
+"Svalöf’s Ugra soybean is a new variety obtained from a cross between a black-seeded American variety, Wisconsin, and a brown-seeded Polish variety, Brunatna Vilenska. " soybean history book
+
+'Ugra' 
+labeled "Ugra", Ugra Saja" & "Ugra Soja"
+Developed by Svalöf AB in Malmöhus, Sweden.
+Pedigree (Wisconsin Black' with 'Brunatna Vilenska' (a Polish variety).)  It was released in 1950 (source:nordgen).
+USDA accession number "Ugra Saja"  is PI 198067 and "Ugra Soja"  at Nordgen NGB 23807
+
+Notes on svalof:
+
+    Svalöf AB later merged with Weibulls AB to form Lantmännen SW seed, a partly Swedish owned seed company.
+The company is headquartered in Landskrona (Southern Sweden).
+The town Landskrona is located within the county administration area called Malmö (once called Malmöhus).
+    "‘Ugra saja’ is actually just named Ugra.
+The ‘saja’ is probably just a misspelling of soja (soya in Swedish)." - Dr.
+Fredrik Fogelberg, Swedish Institute of Agricultural and Environmental Engineering."
+
+\textbf{Altongård}
+
+?
+Holmberg has been generous in sharing his germplasm with other soybean breeders.
+Several Canadian workers have used his lines in their crosses.
+For example, one of the parents of ‘Altona’ is a Holmberg line.?
+
+other text
+
+using the 8 million SNPs of the 409 CCA and 153 SBPA soybeans and the .
+
+To better understand the population relationships of the accessions received from the Nordic gene bank and how they are related to each other, and the rest of the soybean germplasm, they are here compared to a Core Collection of 409 soybeans from the USDA genebank chosen/curated from the soybean germplasm for Central European breeding.
+The Core collection used here is a subset of the soybean germplasm collection, the  USDA genebank accessions, which are assessed to be  preadapted to cultivation in Central Europe, while conserving a high genetic diversity \cite{haupt20}.
+Additionally, accessions were grouped into categories based on the suspected breeding history.
+There are the ten suspected Founders of the Swedish breeding program and the Swedish accessions of the Nordgen genebank.
+
+A Principal Component Analysis (PCA) and the genome-wide Identity By State (IBS) pairwise distance matrix were applied to visualise the overall variance of the data.
+The resulting PCA (Figure \ref{fig:pca}) shows the clustering of the Swedish accessions as a sub-group based on similarity.
+The variance proportions explained.
+
+A closer look at the PCA revealed two accessions of the Nordgen Swedish that were not part of the particular breeding program of interest as seen in (Figure \ref{fig:pca3}  showing the PCA indicating the two accessions not from the breeding program in the cluster with the CCA in orange and black)  and also thereafter confirmed by the genebank passport data that they originated from other Swedish soybean breeding programs (mention more here when certain.
+Altonagård and Ugra soja/ saja, malmåhus).
+This figure also shows the accessions of the Nordgen and core collection that are Swedish of origin and with a name consisting of letters (excluding numbers or non-alphanumeric characters of the accessions' names) as included in the database.
+The IBS (Figure  \ref{fig:dendo}) clearly shows the Swedish accessions of the breeding program of interest clustering together. The  
+ IBS numbers in a table showing ibs within sbpa and out?
+
+ Heatmap based on the sbpa shows high similarity of the SBPA and shows that the 2 accessions Altonagård/Altonagard and Ugra Soja(Saja) are dissimilar.
+
+The population differentiation measured by $F_{\text{ST}}$
+i think we agreed that we needed
+
+change values:
+Fst estimation on genotypes:
+Excluding 0 SNP (monomorphic: TRUE, MAF: NaN, missing rate: NaN)
+    # of samples: 564
+    # of SNPs: 8,533,444
+Method: Weir & Hill, 2002
+### of Populations: 2
+    CCA (409), SBPA (155)
+[1] 0.2412619
+[1] 0.1783931
+     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
+-0.002281  0.048690  0.136954  0.178393  0.263807  0.938701 
+\begin{table}[p]
 \centering 
-\caption{Students and their grades}
+\caption{$F_{\text{ST}}$ estimation on genotypes.
+Measuring the population differentiation of  the 155 Accessions of the Nordgen genebank (Do again with only the 153 that were part of the breeding program)(SBPA) to the Core Collection of 409 accessions (CCA).}
+
 \begin{tableminipage}{\textwidth}
-\begin{tabularx}{\textwidth}{@{}XXXX@{}}
+\begin{tabular}{|l|l|l|l|l|l|l|}
 \hline
-{\bf Student} & {\bf Grade}\footnote{This is an example of a footnote in a table.
-Lowercase, superscript italic letters (a, b, c, etc.) are used by default.
-You can also use *, **, and *** to indicate conventional levels of statistical significance, explained below the table.} & {\bf Rank} & {\bf Notes} \\
-\hline
-Alice & 82\% & 1 & Performed very well.\\
-Bob & 65\% & 3 & Not up to his usual standard.\\
-Charlie & 73\% & 2 & A good attempt.\\
-\hline
-\end{tabularx}
-  \label{tab:shape-functions}
+
+Samples:                        &564        &          &          &          &          &          \\ \hline
+SNPs:                           & 8,533,444 &          &          &          &          &          \\ \hline
+Method: Weir \& Cockerham, 1984 &           &          &          &          &          &          \\ \hline
+Populations:                  & CCA       & 409      & SBPA     & 155      &          &          \\ \hline
+Weighted Fst estimate:          &0.2231884  &          &          &          &          &          \\ \hline
+Mean Fst:                       &0.1108533  &          &          &          &          &          \\ \hline
+summary:                        & Min.      & 1st Qu   & Median   & Mean     & 3rd Qu.  & Max.     \\ \hline
+                                & -0.002515 & 0.011669 & 0.041730 & 0.110853 & 0.145697 & 0.975485 \\ \hline
+\end{tabular}
+  \label{tab:fst}
 \end{tableminipage}
-\end{table*}
+\end{table}
 
-\section{Sample equation}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{Screenshot 2023-07-02 at 14.40.39.png}
+    \caption{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3517337/}
+    \label{fig:enter-label}
+\end{figure}
 
-Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
-\begin{equation}
-S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
-      = \frac{1}{n}\sum_{i}^{n} X_i
-\label{eq:refname1}
-\end{equation}
-denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.
+\begin{table}[p]
+\centering 
+\caption{$F_{\text{ST}}$ estimation on Maturity groups, measuring the population differentiation of  the 155 Accessions of the Nordgen genebank that were part of the breeding program)(SBPA) to the Core Collection of 409 accessions (CCA) groups by the corosponding maturist groups.}
+
+\begin{tableminipage}{\textwidth}
+\begin{tabular}{|l|c|l|l|l|l|l|l|l|l}
+\cline{1-9}
+\multicolumn{1}{|c|}{\begin{tabular}[c]{@{}c@{}}Weir and Cockerham \\ mean Fst estimate\end{tabular}} &                & Group 1                                       & Group 2                           & Group 3                           & Group 4                           & Group 5                           & Group 6                           & Group 7                    &  \\ \cline{1-9}
+                                                                                                      & Maturity group & \multicolumn{1}{c|}{SBPA}\footnote{The Maturity Group of the SBPA are not known and grouped in one group.}\\                     & \multicolumn{1}{c|}{0-0-0}        & \multicolumn{1}{c|}{0-0}          & \multicolumn{1}{c|}{0}            & \multicolumn{1}{c|}{I}            & \multicolumn{1}{c|}{II - IV}      & \multicolumn{1}{c|}{V - X} &  \\ \cline{1-9}
+Group 1                                                                                               & SBPA           & \multicolumn{1}{c|}{\cellcolor[HTML]{E7E6E6}} & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
+Group 2                                                                                               & 0-0-0          & \cellcolor[HTML]{93CC7E}0.2968376             & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
+Group 3                                                                                               & 0-0            & \cellcolor[HTML]{9FD07F}0.2702235             & \cellcolor[HTML]{FBAB77}0.0306470 & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
+Group 4                                                                                               & 0              & \cellcolor[HTML]{B5D680}0.2212166             & \cellcolor[HTML]{FDCD7E}0.0428267 & \cellcolor[HTML]{F98B71}0.0191540 & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
+Group 5                                                                                               & I              & \cellcolor[HTML]{B8D780}0.2146631             & \cellcolor[HTML]{FEE983}0.0525354 & \cellcolor[HTML]{FA9D75}0.0255972 & \cellcolor[HTML]{F8696B}0.0068046 & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
+Group 6                                                                                               & II - IV        & \cellcolor[HTML]{8FCB7E}0.3063683             & \cellcolor[HTML]{FFEB84}0.0532384 & \cellcolor[HTML]{FCC27C}0.0387984 & \cellcolor[HTML]{FBA576}0.0282634 & \cellcolor[HTML]{FA9172}0.0212429 & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
+Group 7                                                                                               & V - X          & \cellcolor[HTML]{63BE7B}0.4046471             & \cellcolor[HTML]{F6E984}0.0745305 & \cellcolor[HTML]{F7E984}0.0728984 & \cellcolor[HTML]{FCEB84}0.0607337 & \cellcolor[HTML]{FFEB84}0.0548296 & \cellcolor[HTML]{FBAA77}0.0301916 & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
+\end{tabular}
+  \label{tab:fst2}
+\end{tableminipage}
+\end{table}
+
+postulate: Population subdivision analysis will quantify the SPBA will be a distinct subpopulation with a moderate to distinct allele frequency difference.
 
@@
 \label{sec:supplementary:material}
-Supplementary Material Methods (S Methods)
-\href{PCA.html}
 
-\section{Data availability}
+(S1: Overview) Detailed descriptions of all supplemental files. 
+(S2: Accessions) Accession and pedigree data 
+(S3: Methods 1) data preparation and filtering 
+(S4: Methods 3) data analysis
+(S5: Results)
+(S6: Historical) Supplementary material historical information
 
-For example: Strains and plasmids are available upon request.
+\subsection{Description of Supplementary Material}
 File S1 contains detailed descriptions of all supplemental files. 
@@
 Gene expression data are available at GEO with the accession number: GDS1234.
-Code used to generate the simulated data can be found at \url{https://github.com/JosephineConnelly/soyadapt_data_analysis}.
+
+\section{Data availability}
+\label{sec:data:availability}
+For example: genotype data wgs are available upon request? 
+File S1 contains detailed descriptions of all supplemental files.
+File S2 contains SNP ID numbers and locations.
+File S3 contains genotypes for each individual.
+Sequence data are available at GenBank and the accession numbers are listed in File S3.
+Gene expression data are available at GEO with the accession number: GDS1234.
+All code can be found at \url{https://github.com/JosephineConnelly/soyadapt_data_analysis}.
 
@@
 
-\section{In-text citations}
-
-Add citations using the \verb|\citep{}| command, for example \citep{neher2013genealogies} or for multiple citations, \citep{neher2013genealogies, rodelsperger2014characterization,Falush16}
+## 
+## To cite R in publications use:
+## 
+##   R Core Team (2021). R: A language and environment for statistical
+##   computing. R Foundation for Statistical Computing, Vienna, Austria.
+##   URL https://www.R-project.org/.
+## 
+## A BibTeX entry for LaTeX users is
+## 
+##   @Manual{,
+##     title = {R: A Language and Environment for Statistical Computing},
+##     author = {{R Core Team}},
+##     organization = {R Foundation for Statistical Computing},
+##     address = {Vienna, Austria},
+##     year = {2021},
+##     url = {https://www.R-project.org/},
+##   }
+## 
+## We have invested a lot of time and effort in creating R, please cite it
+## when using it for data analysis. See also 'citation("pkgname")' for
+## citing R packages.
 
@@
 
+    To insert a bibliography where bibfilename is the name of a .bib file.
+
+\bibliographystyle{genetics}
+
+        To choose a BibTeX bibliographic style file with the extension .bst.
+
 \end{document}

### 2023-08-24 manuscript_soy.tex
24 aug commit

@@
 \cline{1-9}
-\multicolumn{1}{|c|}{\begin{tabular}[c]{@{}c@{}}Weir and Cockerham \\ mean Fst estimate\end{tabular}} &                & Group 1                                       & Group 2                           & Group 3                           & Group 4                           & Group 5                           & Group 6                           & Group 7                    &  \\ \cline{1-9}
-                                                                                                      & Maturity group & \multicolumn{1}{c|}{SBPA}\footnote{The Maturity Group of the SBPA are not known and grouped in one group.}\\                     & \multicolumn{1}{c|}{0-0-0}        & \multicolumn{1}{c|}{0-0}          & \multicolumn{1}{c|}{0}            & \multicolumn{1}{c|}{I}            & \multicolumn{1}{c|}{II - IV}      & \multicolumn{1}{c|}{V - X} &  \\ \cline{1-9}
-Group 1                                                                                               & SBPA           & \multicolumn{1}{c|}{\cellcolor[HTML]{E7E6E6}} & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
-Group 2                                                                                               & 0-0-0          & \cellcolor[HTML]{93CC7E}0.2968376             & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
-Group 3                                                                                               & 0-0            & \cellcolor[HTML]{9FD07F}0.2702235             & \cellcolor[HTML]{FBAB77}0.0306470 & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
-Group 4                                                                                               & 0              & \cellcolor[HTML]{B5D680}0.2212166             & \cellcolor[HTML]{FDCD7E}0.0428267 & \cellcolor[HTML]{F98B71}0.0191540 & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
-Group 5                                                                                               & I              & \cellcolor[HTML]{B8D780}0.2146631             & \cellcolor[HTML]{FEE983}0.0525354 & \cellcolor[HTML]{FA9D75}0.0255972 & \cellcolor[HTML]{F8696B}0.0068046 & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
-Group 6                                                                                               & II - IV        & \cellcolor[HTML]{8FCB7E}0.3063683             & \cellcolor[HTML]{FFEB84}0.0532384 & \cellcolor[HTML]{FCC27C}0.0387984 & \cellcolor[HTML]{FBA576}0.0282634 & \cellcolor[HTML]{FA9172}0.0212429 & \cellcolor[HTML]{E7E6E6}          & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
-Group 7                                                                                               & V - X          & \cellcolor[HTML]{63BE7B}0.4046471             & \cellcolor[HTML]{F6E984}0.0745305 & \cellcolor[HTML]{F7E984}0.0728984 & \cellcolor[HTML]{FCEB84}0.0607337 & \cellcolor[HTML]{FFEB84}0.0548296 & \cellcolor[HTML]{FBAA77}0.0301916 & \cellcolor[HTML]{E7E6E6}   &  \\ \cline{1-9}
+\multicolumn{1}{|c|}{\begin{tabular}[c]{@{}c@{}}Weir and Cockerham \\ mean Fst estimate\end{tabular}} &
+   &
+  Group 1 &
+  Group 2 &
+  Group 3 &
+  Group 4 &
+  Group 5 &
+  Group 6 &
+  Group 7 &
+   \\ \cline{1-9}
+ &
+  Maturity group &
+  \multicolumn{1}{c|}{SBPA} &
+  \multicolumn{1}{c|}{0-0-0} &
+  \multicolumn{1}{c|}{0-0} &
+  \multicolumn{1}{c|}{0} &
+  \multicolumn{1}{c|}{I} &
+  \multicolumn{1}{c|}{II - IV} &
+  \multicolumn{1}{c|}{V - X} &
+   \\ \cline{1-9}
+Group 2 &
+  0-0-0 &
+  \cellcolor[HTML]{93CC7E}0.2968376 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 3 &
+  0-0 &
+  \cellcolor[HTML]{9FD07F}0.2702235 &
+  \cellcolor[HTML]{FBAB77}0.0306470 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 4 &
+  0 &
+  \cellcolor[HTML]{B5D680}0.2212166 &
+  \cellcolor[HTML]{FDCD7E}0.0428267 &
+  \cellcolor[HTML]{F98B71}0.0191540 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 5 &
+  I &
+  \cellcolor[HTML]{B8D780}0.2146631 &
+  \cellcolor[HTML]{FEE983}0.0525354 &
+  \cellcolor[HTML]{FA9D75}0.0255972 &
+  \cellcolor[HTML]{F8696B}0.0068046 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 6 &
+  II - IV &
+  \cellcolor[HTML]{8FCB7E}0.3063683 &
+  \cellcolor[HTML]{FFEB84}0.0532384 &
+  \cellcolor[HTML]{FCC27C}0.0387984 &
+  \cellcolor[HTML]{FBA576}0.0282634 &
+  \cellcolor[HTML]{FA9172}0.0212429 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 7 &
+  V - X &
+  \cellcolor[HTML]{63BE7B}0.4046471 &
+  \cellcolor[HTML]{F6E984}0.0745305 &
+  \cellcolor[HTML]{F7E984}0.0728984 &
+  \cellcolor[HTML]{FCEB84}0.0607337 &
+  \cellcolor[HTML]{FFEB84}0.0548296 &
+  \cellcolor[HTML]{FBAA77}0.0301916 &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
 \end{tabular}

### 2023-09-21 manuscript_soy.tex
21 sep

@@
- \documentclass[9pt, onecolumn,twoside]{gsajnl}
+ \documentclass[9pt, twocolumn,twoside]{gsajnl}
 % how i want it for inport into word: \documentclass[9pt, onecolumn,twoside]{gsajnl}
@@
 \usepackage{epstopdf}
+\usepackage{wrapfig}
 \usepackage[T1]{fontenc}
@@
 
-\title{SoyAdapt}
+\title{Soybeans for the Future: Population Genetic Analysis Of A Historical Swedish Soybean Population For Future Germpasm Utalisation}
+
+%Swedish Soybean Genetic Diversity: Influence of Breeding Practices and Identification of Selection Signatures
+
+%Genetic Diversity in a Swedish Soybean Population: Signatures of Selection and Implications for Future Soybean Breeding in Northern Environments
+
+% Swedish Soybean Genetic Diversity: Influence of Breeding Practices and Identification of Selection Signatures
+
+%Soybean Genetic Variation in Swedish population: A Study of Breeding Influence and Selection Signatures
+
+%Swedish Soybean Genetic Diversity: Influence of Breeding Practices and Identification of Selection Signatures
+
+%Genetic Diversity in the Swedish Soybean Population: Impacts of the Breeding Program, Selection Signatures, and Implications for Future Germplasm Utilization
+
+%Population Genetic Insights into the Swedish Soybean Population: Impact of Breeding Strategies and Selection Signatures
 
@@
 \begin{abstract}
-Soybean (Glycine max L. Merrill) is the worlds leading  oilseed crop used as a  primary source of vegetable oil for human consumption and in protein meal for animal feed.
+Soybean (Glycine max L. Merrill) is the world's leading oilseed crop used as a  primary source of vegetable oil for human consumption and protein meal for animal feed. 
+A breeding program of the 1940s-1970s resulted in 160 accessions later donated to the seed bank of the Nordic Genetic Resource Center (NordGen).
+Climate change and plant protein independence have driven the idea of trying to repeat the effort of plant genetic enhancement for soybean cultivation in northwest Europe.
+Uncovering the population genetics of this as yet unutilised germplasm can give a head start to these future soybean breeding efforts.
+The Swedish soybeans are revealed as a population that has undergone bottlenecks, resulting in reduced effective population size (Ne), elevated levels of genetic drift, decreased nucleotide diversity, and increased linkage, all attributed to the selective breeding efforts or the selection of the genetic progenitors.
 
-Search engine optimization (SEO)
+The Core collection (CCA) was created by \cite{haupt20}  as a collection identified from the USDA germplasm for central European soybean breeding.
+Comparison of the CCA to the Swedish breeding program accessions (SBPA) has been used throughout the paper to put the SBPA into the perspective of germplasm likely to be utilised in these future soybean breeding efforts.
+The SBPA has half the nucleotide diversity of the CCA.
+Fst analysis for population subdivision reveals a population difference of 0.22 calculated for a weighted Fst estimate.
 
-There are a few simple ways to maximize your article's discoverability and search results.
+genomic regions of selection differentiation were detected between the CCA and the SBPA by using snp specific genomewide pairwise Fst and to uncover
 
-    Include a few of your article's keywords in the title of the article
+The reduced level of genetic variation in elite breeding populations means that in the case of soybean, there is a need to increase germplasm utilization for abiotic adaptation to the colder conditions of Northern Europe.
+This research means to examine the genetic structure and diversity within a population of soybeans originating from a breeding program during the 1940s to 1970s in Sweden (SBPA).
+We aim to gain insights into the pressures that have shaped the genetic makeup of these soybean accessions.
+So that this resource of untold potential can be fully utilized.
 
-    Do not use long article titles
-
-    Pick 5-8 keywords using a mix of generic and more specific terms on the article subject(s)
-
-    Use the maximum amount of keywords in the first two sentences of the abstract
-
-    Use some of the keywords in level 1 headings
-
-    Title
-
-The title should be concise, omitting terms that are implicit and, where possible, be a statement of the main result or conclusion presented in the manuscript.
-Abbreviations should be avoided within the title.
-
-Witty or creative titles are welcome, but only if relevant and within measure.
-Consider if a title meant to be thought-provoking might be misinterpreted as offensive or alarming.
-In extreme cases, the editorial office may veto a title and propose an alternative.
-
-Authors should avoid:
-
-    titles that are a mere question without giving the answer
-
-    unambitious titles, for example starting with 'Towards,' 'A description of,' 'A characterization of' or 'Preliminary study on'
-
-    vague titles, for example starting with 'Role of', 'Link between', or 'Effect of' that do not specify the role, link, or effect
-
-    including terms that are out of place, for example, the taxonomic affiliation apart from species name.
-
-For Corrigenda, General Commentaries, and Editorials, the title of your manuscript should have the following format.
-
-    'Corrigendum: [Title of original article]'
-
-    General Commentaries:
-    'Commentary: [Title of original article]'
-    'Response: Commentary: [Title of original article]'
-
-    'Editorial: [Title of Research Topic]'
 \end{abstract}
 
-\keywords{Soy; population genetics; Diversity; Selection; more keywords}
+\keywords{Soybean; Glycine max; Population Genetics; Diversity; Fst; Linkage Disequalibrium; Selection Signatures; Plant Breeding}
 
-\dates{\rec{01 09, 2023} \acc{01 09, 2023}}
+\dates{\rec{29 09, 2023} \acc{29 09, 2023}}
 
@@
 
-\textit{what we already know:} Several severe genetic bottlenecks occurred in soybean.
+This is a link to part A where there is the extensive introduction/litterature review:
+https://www.overleaf.com/2689273778mwyzpfmjvbmf
+
+ps abrtract is in the preamble...
+
+Soybean is Glycine max, this and that and important.
+
+Several severe genetic bottlenecks occurred in soybean during domestication and the continuous breeding efforts.
 Compared to the wild species, the genetic diversity was halved, resulting in a loss of 81\% of rare alleles (Hyten et al., 2006).
-Additionally, two major bottlenecks occurred during the development of North American modern cultivars, where only a few landraces were used.
-And as a result of intensive soybean breeding over the past 75 years.
+Additionally, two major bottlenecks occurred during the development of North American modern cultivars, where only a few landraces were used and as a result of intensive soybean breeding over the past 75 years.
 Elite cultivars have emerged, but they are derived from only about 19 landraces.
-Consequently, the North American breeding pools now retain only 72\% of genome diversity and have lost 79\% of rare alleles found in diverse landraces (Gizlice et al., 1996).
+Consequently, the North American breeding pools now retain only 72\% of genome diversity and have lost 79\% of rare alleles found in diverse landraces \cite{gizlice96}.
 
-\textit{soybean needs to be grown in the north}:
-Modern soybean breeding has led to yield increases of $229 \, \text{kg} \, ha^{-1} \, \text{yr}^{-1}$ (Rincker et al., 2014), gaining a deeper understanding of the genomic basis of this progress could offer valuable insights for advancing and adapting soybean cultivation even further.
+soybean needs to be grown in the north:
 
-Expanding the production of soy to the northern climates by leveraging the advancements in soybean breeding and focusing on the genes relevant to adaption, can diversify the global supply chain, reduce dependency, and potentially alleviate the environmental pressures associated with soy production in high biodiversity, sensitive areas/ecologically vulnerable regions.
-With the global demand for soy and the need for sustainable agricultural practices, investing in breeding programs to develop soybean varieties suitable for northern climates can present a promising solution.
+With the European Green Deal of 2019, there is an intention of harnessing the potential of EU plant protein production.
+This intention is motivated as a response to the needs of farmers, producers and consumers.
+The production and sourcing of plant-based proteins in the EU also carries with it a range of environmental benefits such as a decrease in imports and therefore transport pollution reduction.
+
+The difficulties identified by the Report: Development of plant proteins in the European Union \cite{EC18} are:
+
+- the agronomic conditions in Europe, which are not optimal for large-scale production of plant proteins; 
+- the economic profitability of these crops in Europe; 
+- the competitiveness of EU protein crops compared to imported plant proteins; 
+- competition over the use of arable land; 
+- and a lack of research on breeding, agronomic practices and different uses.
+
+(Danish field trials have shown the Swedish cultivar Fiskeby V to yield only 11.2hkg/ha \cite{petersen09,petersen10}.
+Modern soybean breeding has led to yield increases of $229 \, \text{kg} \, ha^{-1} \, \text{yr}^{-1}$ \cite{rincker14}, advancing and adapting soybean cultivation even further.
+)
+
+Expanding the production of soy to the northern climates by leveraging the advancements in soybean breeding and focusing on the genes relevant to adaptation can diversify the global supply chain, reduce dependency, and potentially alleviate the environmental pressures associated with soy production in high biodiversity, sensitive areas/ecologically vulnerable regions.
+With the global demand for soy and the need for sustainable agricultural practices, investing in breeding programs to develop soybean varieties suitable for northern climates can present a promising part of the solution.
 
 \textit{the Swedish soybeans (goals and scope)}
+The Swedish breeding population SBPA consists of the soybean germplasm collection from the seed bank of the Nordic Genetic Resource Center (NordGen).
+One of the cultivars that resulted from this breeding program is well-known as a genetic source of
+goals
+not technical
 
-\textit{mention approaches}
+scope
 
-\textit{In this study we}
-The goal is to help precipitate the utilisation of these soybean genetic resources. to increase germplasm utilization for through understanding what we have in the genebanks.
+\textit{In this study}
+
+In this study, we present insights into the population genetic analysis of the soybean accessions obtained from the NorGen genebank accessions. 
+There haven't been any past investigations into this germplasm resource, so therefore, the goal of this project was to better understand the value of the genebank collection so as to help precipitate the utilisation of these soybean genetic resources.
+to increase germplasm utilization through understanding what we have in the genebanks.
+We present population stratification, diversity and linkage of the population.
+Through population genetic analyses supported by breeders publications and historical documents, we can infer what happened and what  to this population and determine the genetic significance of the resource.
+Presented is a comprehensive population genetic analysis of the SBPA soybean population. Determining the genetic diversity and population structure of the SBPA.
+An assessment of how the SBPA compare to the broader genetic diversity in the species through a comparative analysis with a core collection (CCA).
+And a test of selection efficacy of the Swedish breeding program by screening biological pathways.
 
@@
 155 soybean accessions were obtained from the Nordgen genebank of Swedish origin.
-These accessions originate from a Swedish breeding program (SBP) running from the 1840s to the 1970s which used material consisting of a mixture of elite cultivars of the time and Japanese landraces, breeding for the adaptation to the cool climate of northwestern Europe \cite{holmberg1973}.
+These accessions originate from a Swedish breeding program (SBP) running from the 1840s to the 1970s, which used material consisting of a mixture of elite cultivars of the time and Japanese/Siberian landraces, breeding for the adaptation to the cool climate of northwestern Europe \cite{holmberg1973}.
+The Swedish breeding program (SBP) of the 1940s-1980s of Algot Holmberg and Sons from Norrkoping produced the known cultivars Fiskeby III, Fiskeby V, Träff, and Bråvalla, which are included in the NordGen gene bank accessions and in this analysis excluding Fiskeby III that was not available from the Nordgen genebank at the time.
 
@@
 The Core collection accessions (CCA) used here is a subset of the large soybean germplasm collection, the  USDA genebank accessions consisting of 415 of the USDA soy germplasm accessions.
-This Core collection has been selected with a focus on adaptation to high-latitude cold regions using environmental data from phenotypic trials in Germany, and comparing Donor opulation of Environments (DPE) in Asia and the Target Population of Environments (TPE) in Central Europe by \cite{haupt20}.
+This Core collection has been selected with a focus on adaptation to high-latitude cold regions using environmental data from phenotypic trials in Germany and comparing Donor opulation of Environments (DPE) in Asia and the Target Population of Environments (TPE) in Central Europe by \cite{haupt20}.
 From the more than 17.000 accessions, two diverse core collections of 183 and 366 accessions were created.
-These 514 diversity panels are used here due to that they are likely preadapted to cultivation in Central Europe, while simultaneously conserving a high level of genetic diversity.
+These 514 diversity panels are used here due to that they are likely preadapted to cultivation in Central Europe while simultaneously conserving a high level of genetic diversity.
 
@@
 These accessions have been previously genotyped with the SoySNP50K iSelect BeadChip  \cite{song13,song15} has been used for genotyping the USDA Soybean Germplasm Collection and is available for use from SoyBase (https://soybase.org/) in the version mapped to the soybean genome assembly Glyma.Wm82.a2  \cite{schmutz10}.
-In the pedigree information from Nordgen SBPA, are 10 accessions present in the USDA genebank and thus are part of the SoySNP50K genotypic data.
+In the pedigree information from Nordgen SBPA, there are 10 accessions present in the USDA genebank and thus are part of the SoySNP50K genotypic data.
 These specific accessions were included as they are considered Founders of the SBP.
@@
 Of the Nordgen accessions, Four were of Latvian origin and therefore excluded in further analysis.
-The reads were aligned to the Williams 82 2a reference genome \cite{schmutz10} using an adapted GATK Best Practices \cite{mckenna10}, keeping MQ30 and biallelic sites and using VCFTools \cite{danecek11} applying a minor allele frequency (MAF) filter of 0.01 in order to remove rare variants from the data, reducing the number of SNPs markers were initially called (raw dataset) from 17,648,123 to 10,000,122.
+The reads were aligned to the Williams 82 2a reference genome \cite{schmutz10} using an adapted GATK Best Practices \cite{mckenna10}, keeping MQ30 and biallelic sites and using VCFtools v.
+0.1.16 \cite{danecek11} applying a minor allele frequency (MAF) filter of 0.01 in order to remove rare variants from the data, reducing the number of SNPs markers were initially called (raw dataset) from 17,648,123 to 10,000,122.
 For a detailed account of accessions, programs and commands used see supplementary material: \nameref{sec:supplementary:material} 
@@
 This data set was used to compare the Allele Frequency of the two populations.
-For all other analyses, MAF 5\%  was removed using Vcftools (--maf 0.05) on the dataset as a whole, except for the case of LD calculation MAF5\% was removed separately for SBPA and CCA.
+For all other analyses, MAF 5\%  was removed using VCFtools (--maf 0.05) on the dataset as a whole, except for the case of LD calculation MAF5\% was removed separately for SBPA and CCA.
 
@@
 
+* here insert a table with information of the snps and stuff.
+
 \subsection{Population Structure Analysis}
@@
 The package SNPRelate in R was used to compute an identity-by-state (IBS) matrix that calculates the proportion that two randomly selected reads that contain a certain SNP locus are the same or different between two individuals.
-The resulting pairwise IBS matrix was used to generate a cluster dendrogram using the function \textit{snpgdsHCluster: Hierarchical cluster analysis In SNPRelate }in R.
+The resulting pairwise IBS matrix was used to conduct a hierarchical cluster analysis to generate a dendrogram using the function \textit{snpgdsHCluster: Hierarchical cluster analysis In SNPRelate }in R.
 
+F-statistics, particularly Fst, quantifies the distribution of genetic diversity by measuring the variation in allele frequencies among different populations. 
 Relatedness measures F-statistics (also known as fixation indices) for given populations were calculated using the function \textit{snpgdsFst: F-statistics (fixation indices) in SNPRelate }in R.
@@
 
-Later in the analysis when looking for signatures of selection Fst is calculated according to the methods of \cite{Weir84} using the Vcftools \cite{danecek11} function --weir-fst-pop on the SNP variants with (no window size) a window size of 1 Mb with a step size of 100 kb.
-
-\subsection{Allele frequency, LD and nucleotide diversity}
+\subsection{Genetic Diversity Estimates}
 The idea is to look into the effects of the bottleneck caused by the SBP, the increased drift and will be detected as a reduction of Ne.
@@
 Allele Frequency (AF) was calculated using Vcftools --freq and plotted in R.
-For the LD analysis vcftools --maf 0.05 for each population was removed separately.
-The function \textit{snpgdsLDmat: Linkage Disequilibrium (LD) analysis, in SNPRelate } in R.
-A matrix of "r" for R coefficient (by EM algorithm assuming HWE) is calculated for pairwise snps separately for each chromsome.
+LD analysis was preformed in VCFtools  --maf 0.05 for each population was removed separately.
+The function \textit{snpgdsLDmat: Linkage Disequilibrium (LD) analysis, in SNPRelate }in R.
+A matrix of "r" for R coefficient (by EM algorithm assuming HWE) is calculated for pairwise SNPs separately for each chromosome.
 The squared correlation coefficient r2 is a popular measure of LD an chosen due to the advantage of this calculation being less sensitive to marginal allele frequencies over  D or D' \cite{sved18}.
@@
 The $r^2$  values were binned and plotted using ggplot2 \cite{wickham16} in R (ref).
-Nucleotide diversity ($\pi$), measuring the degree of polymorphism in the populations was calculated with the wgs data using vcftools for 100 kb windows --window-pi 10000 \cite{nei79}.
-*This is appropriate @ 25kb as we just discussed earlier today with the fst.
+Nucleotide diversity ($\pi$), measuring the degree of polymorphism in the populations, was calculated with the wgs data using VCFtools for 100 kb windows --window-pi 10000 \cite{nei79}.
 
-\subsection{H3}
-Q and Method
+\subsection{Selection Signature Detection }
+For testing signatures of selection, TajimasD statistic and pairwise FstFst is calculated according to the methods of \cite{Weir84} using the Vcftools function --weir-fst-pop on the SNP variants with no sliding window, but a window size of 25kbp --fst-window-size 25000.
+The same window is used for calculating TajimasD, --TajimaD 25000.
+This window-based approach for Fst statistics has proven successful in minimizing sampling error while preserving the genuine signal when investigating evidence of selection in populations \cite{beissinger15}.
+
+Candidate gene enrichment analysis. 
+Pairwise Fst values across the same window were used to assess the divergence between the SBPA and the CCA and Tajima's D for each population was calculated over the same window. 
+gff3 annotation of the genome data was used to further analyse whether the higher Fst values or low Tajima's D values are significant for genes correlated to photoperiod or cold tolerance traits. 
+Pathway screening analysis.
+To look for enrichment or degradation of genes in other pathways enrichment was tested.
+
+GeneIds, locations and GO functions were downloaded via Soybase.org and genes were selected based on a word search of the Gene Ontology biological process descriptions for flowering and cold tolerance.
+A further 14 genes were selected for enrichment analysis based on genes known to be involved in photoperiod regulation.
+Enrichment analysis was used to see if there was a significant difference by testing the significance of the number of genes above or below an Fst and TajimasD threshold.
+For the Fst threshold, the upper five per cent value was used and for Tajimas D, any value under zero.
+For details on the analysis, see the link to the R code in the section on data availability.
+\label{sec:data:availability}
 
@@
 
-Indicate what statistical analysis has been performed.
+Fisher's exact test was chosen to test significance of enrichment because Chi squared can be unreliable if some contingency tables have cells of low numbers.
+This is the case in the pathway screening analysis, and alternatively the Fisher's exact test can be used when there are low expected values.
+The test gives the exact probability of the observed values \cite{quicke21}.
+Fisher's exact test was used to test for the pathway screening analysis (the data are counts comparing variables) to test whether there is a disparity between the variables, frequencies of genes above and below the chosen threshold of Fst and TajimasD of genes within or not within the pathway genes.
 
-\section{Results and Discussion}
+Caveats of the statistical test. 
+Replicates? Fst is calculated on the 160 samples of The SBPA as a population. but if there is a very high degree of similarity in the population, there are effective pseudo-replicates of the samples.
 
-\subsection{H1 The Swedish soybean accessions were part of the SBP}
-The accessions from the NordGen gene bank (Swedish breeding program accessions, SBPA) are closely related and therefore we can conclude that they were part of the same breeding program, the Swedish breeding program (SBP), of the 1940s-1980s of Algot Holmberg and Sons from Norrköping, producing the known cultivars, Fiskeby III, Fiskeby V, Träff $\&$ Bråvalla, which are included in the NordGen gene bank accessions and in this analysis.
-Two accessions are identified as accessions that were not part of the Swedish breeding program of interest but have been part of other Swedish or Nordic soybean breeding trajectories and later obtained from the Russian seed bank.
+\section{Results}
 
-\subsubsection{Relatedness: The SBPA are genetically similar and therefore closely related}
-There are 155 soybean accessions sequenced from the Nordgen genebank.
-153 of the accessions are similar to a degree that we can conclude they are closely related to a degree of XXX.
-numbers from the ibs?
-Two of the Swedish accessions (Altonagård \& Ugra Soja) show up clearly distinct from the cluster of SBPA and are therefore no longer considered or included as accessions of the investigated (SBP).  
-\begin{figure}
+The Swedish accessions are of a cohesive unit, the same breeding program; they are genetically similar, and there are clear indications that the population has undergone a recent bottleneck, that leads us to infer that the is a small effective population size (Ne).
+Effective population size determines genetic variation, genetic drift and linkage disequilibrium in populations \cite{tenesa07}.
+The results of these effects are seen to indicate drift and low genetic variation, a high number of fixed alleles and high LD decay.
+Despite strong drift, a screening of the SBPA has resulted in evidence of enrichment in 34 pathways and evidence of a degradation in three pathways.
+
+\subsubsection{Relatedness: The SBPA are genetically similar and, therefore closely related}
+The SBPA are closely related, and therefore, we can conclude that they were part of the same breeding program.
+This is supported by three sources of metadata: firstly, the names of the accessions align with the breeders' coding of crosses; secondly, the seeds were contributed by the breeding company; and thirdly, the pedigree information further substantiates this finding \cite{holmberg1973}.
+
+In the initial analysis, we discovered two accessions, namely Altonagård and Ugra Soja, which, while not originally associated with the specific Swedish breeding program of interest, have significant historical ties to other Swedish and Nordic soybean breeding initiatives.
+These accessions were subsequently acquired by the Nordgen genebank through a Russian genebank.
+It's worth noting that these two accessions do not align with the SBPA in either the hierarchical cluster dendrogram where we consider pairwise identity-by-state (IBS) analysis or the PCA analysis, as illustrated in a PCA of the accessions aquired from the Nordgen genebank, see \nameref{sec:supplementary:material} S5.1
+
+Figure \ref{fig:pca}.
+shows two main clusters, the SBPA clustered together and the larger CCA clustered above.
+The accessions identified as the Founders of the breeding program are situated between and within the two clusters.
+Figure \ref{fig:pca}.
+shows the first two principal components (PCs) of the PCA, comprising 10.3 and 3.4\% of the variation, respectively.
+The PCA captures 22.82\% of the variation in the first 10 PCs where combinations of the first 4 can be seen in the Supplementary Material figure S5.2.
+It is clear that the SBPA clusters together on all 6 combinations of the first 4 PCs.
+The high sample number of the SBPA (153) along with the high similarity drags the SBPA cluster away from the larger cluster consisting of the CCA.
+The importance of the sample size for distance in PC1 is demonstrated in S5.2 where a smaller sample size of the SBPA is used.
+Here it is clear that although the SBPA accessions are not in a cluster at a distance from the rest, the SBPA still cluster together in all of the first 4 PCs explaining together 12.8\% of the variance, confirming the SBPA similarity.
+
+The Hierarchical cluster analysis based on pairwise identity-by-state (IBS) values from the wgs SNPs also indicates the SBPA as a group with high similarity.  *more
+
+\subsubsection{Distinctness: The SBPA are a genetically cohesive group}
+The SBPA are a genetically cohesive group of accessions, as they cluster together and are removed from the CCA as a whole.
+Figure IBS heatmap shows the SBPA clustered together in the top left-hand corner of the heatmap.
+The clustering is made from the IBS calculated similarity, and the hierarchical cluster analysis shows the SBPA clustering together within the group.
+This clustering has a very high similarity to the  pedigree dendrogram S2.3.
+Figure PCA shows the SBPA as clustered together by the first principal component, capturing 10.3\% of the variation.
+
+It is also clear that there are accessions of the CCA that are closely related to the SBPA,.
+This is expected from a domesticated crop such as soybean because of the known pedigrees and breeding programs exchanging germplasm.
+In the CCA, 94 of the accessions are of European origin.
+Additionally, seven of those originate from Sweden, some with accession names are identical or similar to that of the SBPA.
+See S2 for all accessions and  S5.1 for a PCA showing accessions with names in both datasets.
+A few of the CCA accessions are also thought to be a Founder of the SBPA.
+
+\subsubsection{Divergence: There is a divergence within the subpopulation SBPA of the soybean when comparing other groups of soybean}
+Population subdivision analysis will quantifys the SPBA as a distinct subpopulation with a distinct allele frequency difference.
+
+The result of divergent selection and drift, causes genetic differentiation in separate populations over time.  
+The SBPA are genetically different to the CCA, Fst was calculated using SBPA as a subpopulation of the CCA. The CCA has a variability comparable to the soybean germplasm as a whole \cite{haupt20}. 
+The Weighted Fst estimate is  0.22, and the Mean Fst 0.11 between SBPA as a subpopulation of CCA, indicating a high level of genetic differentiation between the populations (Figure FST1).  
+percent reduction in heterozygosity in the total population due to population structuring
+
+To further quantify population differentiation, the CCA was grouped into subpopulations based on Maturation score status to compare the divergence of the SBPA where no maturation scores have been given Table  \ref{tab:fst}.
+The high Fst of the SBPA shows a considerable degree of differentiation compared to the other groups defined by the maturity group.
+A relatively high Fst mean between (0.405 \& 0.215) for the SBPA compared to the CCA groups shows a considerable degree of differentiation in relation to the other populations as defined by the maturity group.
+The CCA in relation to each other has a relatively low Fst, between (0.075 \& 0.007), showing that differentiation is not high when grouping the CCA by Maturity.
+
+\subsubsection{Founders}
+A list of known and suspected founders has been included in the figure PCA.
+It is clear that some of the founders are closely related to the SBPA as they are situated between the SBPA and CCA clusters.
+This corroborates the idea of these accessions being founders of the SBP.
+Additionally, accessions of the founders cluster on the IBS hierarchical analysis close to the SBPA (S5.x).
+They are also in the PCs situated close to the SBPA.
+Other founder accessions are not distinguishable as Founders in this way.
+There are also CCA accessions that have been found to actually be SBPA.
+all this is noted in the large IBS dendrogram of the accessions and in the S2 accessions names and annotations  It is also expected that a founder accession could be more related to accessions in the CCA than the SBPA because of these accessions use in other breeding tajectories.
+A more detailed picture of decent can be seen in the supplementary material where the pedigree data of the SBPA are collected.
+\nameref{sec:supplementary:material}.
+S2.3.
+
+\begin{figure}[p]
 \centering
-\includegraphics[width=\linewidth]{plot_heatmap.pdf}
-\caption{This heatmap includes Altonagård and Ugra soja which are seen as the bottom two blue acessions. They are dissimilar to eachother and dissimmilar to the rest of the accessions.}
-\label{fig:heat}
+\includegraphics[width=\linewidth]{PCA1.pdf}
+\caption{Biplot of the principal component analysis of soybean genotypes.
+The colours represent the defined origin groups, and the shapes indicate whether the accessions are from the Core Collection Accessions (CCA), the Swedish Breeding Program Accessions (SBPA) or the Founders of the breeding program (F).}
+\label{fig:pca}
 \end{figure}
 
-Characterise genetic relatedness. 
-PCA the first 10 pcs capture 22.82\%  of the variation.
-the ibs shows identity towards 1 to be at: numbers?
-
-\subsubsection{Distinctness: The SBPA are a genetically cohesive group}
-The SBPA are a genetically cohesive group of accessions.
-In comparison to the CCA, they are similar to each other (matrix numbers?) as seen in the figure the SBPA cluster within the dendrogram of the CCA and the SBPA.
-It is also clear that some of the CCA are closely related to the SBPA, which is expected as a domesticated crop and the history known of other pedigrees and breeding programs exchanging germplasm.
-In the CCA, 94 of the accessions are of European origin and additionally, seven of those  originate from Sweden, some with accession names identical or similar to that of the SBPA.
-\begin{figure}
+\begin{figure}[p]
+\centering
+\includegraphics[width=\linewidth]{heat1.pdf}
+\caption{This heatmap is a visual representation of the IBS values of the SBPA and CCA accessions.
+IBS values that are from 0 to 1 where 1 is identical by state.
+In the upper left corner are the SBPA accessions collected by the IBS matrix and hierarchical clustering-based analysis.
+The SBPA IBS values range from 0.73 to 0.97 for the SBPA and 0.56 to 0.98 for the CCA.
+The dataset as a whole, therefore ranges from 0.56 to 0.98.
+}
+\label{fig:heatall}
+\end{figure}
+\begin{figure}[p]
 \centering
@@
 
-\subsubsection{Divergence: There is a (significant) divergence within the subpopulation SBPA of the soybean when comparing other groups of soybean}
+\begin{wraptable}{r}{0.5\textwidth}
+\begin{table}[p]
+\centering 
+\caption{$F_{\text{ST}}$ estimation on genotypes.
+Measuring the population differentiation of  the 155 Accessions of the Nordgen genebank (Do again with only the 153 that were part of the breeding program)(SBPA) to the Core Collection of 409 accessions (CCA).}
 
-Genetic differentiation, e.g., Fst was calculated using SBPA as a subpopulation of the CCA.
-This is argumentatively valid since the CCA has a variability comparable to the soybean germplasm as a whole (ref).
-Fst mean at 0.178
-The CCA has furthermore been grouped into subpopulations according to Maturation score status to compare the divergence of the SBPA. 
-Figure: Fst
-The high Fst of the SBPA shows that there is a considerable degree of differentiation in comparison to the other groups defined by maturity group shows
+\begin{tableminipage}{\textwidth}
+\begin{tabular}{|l|l|l|l|l|l|l|}
+\hline
 
-A relatively high Fst mean between (0.405 \& 0.215) for the SBPA in comparison to the CCA groups shows that there is a considerable degree of differentiation in relation to the other populations as defined by the maturity group.
-The CCA in relation to each other has a relatively low Fst, between (0.075 \& 0.007) showing that differentiation is not high when grouping the CCA by Maturity.
+Samples:                        &564        &          &          &          &          &          \\ \hline
+SNPs:                           & 8,533,444 &          &          &          &          &          \\ \hline
+Method: Weir \& Cockerham, 1984 &           &          &          &          &          &          \\ \hline
+Populations:                  & CCA       & 409      & SBPA     & 155      &          &          \\ \hline
+Weighted Fst estimate:          &0.2231884  &          &          &          &          &          \\ \hline
+Mean Fst:                       &0.1108533  &          &          &          &          &          \\ \hline
+summary:                        & Min.      & 1st Qu   & Median   & Mean     & 3rd Qu.  & Max.     \\ \hline
+                                & -0.002515 & 0.011669 & 0.041730 & 0.110853 & 0.145697 & 0.975485 \\ \hline
+\end{tabular}
+  \label{tab:fst}
+\end{tableminipage}
+\end{table}
+\end{wraptable}
+\begin{wraptable}{r}{0.5\textwidth}
+\begin{table}
+\centering 
+\caption{$F_{\text{ST}}$ estimation on Maturity groups, measuring the population differentiation of  the 155 Accessions of the Nordgen genebank that were part of the breeding program)(SBPA) to the Core Collection of 409 accessions (CCA) groups by the corrosponding maturity groups.}
 
-\textbf{Founders}
-p4. summary of what we can see of the founders. 
-A list of known and suspected founders has been included in the PCA.
-It is clear that (some specific) founders are closely related to the SBPA and were clearly founders of the SBP whereas others are not distinguishable as Founders.
-It is expected that the founder accessions could be more related to accessions in the CCA than the SBPA.
-A more detailed picture of decent can be seen in the supplementary material where the pedigree data of the SBPA are collected.
-\nameref{sec:supplementary:material}.
+\begin{tableminipage}{\textwidth}
+\begin{tabular}{|l|c|l|l|l|l|l|l|l|l}
+\cline{1-9}
+\multicolumn{1}{|c|}{\begin{tabular}[c]{@{}c@{}}Weir and Cockerham \\ mean Fst estimate\end{tabular}} &
+   &
+  Group 1 &
+  Group 2 &
+  Group 3 &
+  Group 4 &
+  Group 5 &
+  Group 6 &
+  Group 7 &
+   \\ \cline{1-9}
+ &
+  Maturity group &
+  \multicolumn{1}{c|}{SBPA} &
+  \multicolumn{1}{c|}{0-0-0} &
+  \multicolumn{1}{c|}{0-0} &
+  \multicolumn{1}{c|}{0} &
+  \multicolumn{1}{c|}{I} &
+  \multicolumn{1}{c|}{II - IV} &
+  \multicolumn{1}{c|}{V - X} &
+   \\ \cline{1-9}
+Group 2 &
+  0-0-0 &
+  \cellcolor[HTML]{93CC7E}0.2968376 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 3 &
+  0-0 &
+  \cellcolor[HTML]{9FD07F}0.2702235 &
+  \cellcolor[HTML]{FBAB77}0.0306470 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 4 &
+  0 &
+  \cellcolor[HTML]{B5D680}0.2212166 &
+  \cellcolor[HTML]{FDCD7E}0.0428267 &
+  \cellcolor[HTML]{F98B71}0.0191540 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 5 &
+  I &
+  \cellcolor[HTML]{B8D780}0.2146631 &
+  \cellcolor[HTML]{FEE983}0.0525354 &
+  \cellcolor[HTML]{FA9D75}0.0255972 &
+  \cellcolor[HTML]{F8696B}0.0068046 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 6 &
+  II - IV &
+  \cellcolor[HTML]{8FCB7E}0.3063683 &
+  \cellcolor[HTML]{FFEB84}0.0532384 &
+  \cellcolor[HTML]{FCC27C}0.0387984 &
+  \cellcolor[HTML]{FBA576}0.0282634 &
+  \cellcolor[HTML]{FA9172}0.0212429 &
+  \cellcolor[HTML]{E7E6E6} &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+Group 7 &
+  V - X &
+  \cellcolor[HTML]{63BE7B}0.4046471 &
+  \cellcolor[HTML]{F6E984}0.0745305 &
+  \cellcolor[HTML]{F7E984}0.0728984 &
+  \cellcolor[HTML]{FCEB84}0.0607337 &
+  \cellcolor[HTML]{FFEB84}0.0548296 &
+  \cellcolor[HTML]{FBAA77}0.0301916 &
+  \cellcolor[HTML]{E7E6E6} &
+   \\ \cline{1-9}
+\end{tabular}
+  \label{tab:fst2}
+\end{tableminipage}
+\end{table}
+\end{wraptable}
 
-Figure: PCA with founders in. 
-The founders could also be indicated earlier in the dendrogram clustering together with the SBPA and the SBPA cluster nested within the founders.
+\subsection{H2 Low Genetic diversity and high LD decay in the SBPA point to a bottleneck and low Ne} 
+There is a reduction in effective population size detected in a high level of linkage of the SBPA and a lower nucleotide diversity ($\pi$ ) in comparison to the CCA.
+There is also an accumulation of rare alleles and fixed alleles in the SBPA population.
+These observations suggest a bottleneck making an impact on the genetic composition of the SBPA population ensued by the intensification of genetic drift.
 
-\subsection{H2 Genetic diversity}
+\subsubsection{The SBP caused a bottleneck as seen in a decrease in genetic diversity }
+Earlier studies have shown a halving in nucleotide diversity when comparing wild soybean to their domesticated progenitors and a further halving in nucleotide diversity in breeding regimes in recent years comparing landraces to elite cultivars \cite{kim21}.
+The SBPA have a mean nucleotide diversity of ($\pi$ = 0.000999), or a difference of  9/10.000 bases between two sequences.
+Whereas the CCA has almost double the mean nucleotide diversity of  ($\pi$ = 0.001911) or 19/10.000 bases differing.
+The CCA has two times higher nucleotide diversity than the SBPA.
+This fall in nucleotide diversity is consistent with a bottleneck during the selection of founders and subsequent selection by the SBP.
 
-\textbf{The SBP caused a bottleneck as seen in a decrease in genetic diversity }
+\subsubsection{Allele Frequency shows a high level of drift}
+A high number of alleles at high frequency shows that the SBPA have a higher number of monomorphic locus than the CCA.
+A high number of fixed alleles together with the low nucleotide diversity can arise through natural selection, genetic bottlenecks, or genetic isolation in the population.
+The SBPA also has a higher number of genetic variants at low frequencies than the CCA Figure AF.
+Rare alleles arise through mutation, genetic drift, or migration from other populations.
+A high number of allele frequencies in the extremities of the frequency spectrum is a pattern of increased allel frequencies leading to fixation due to drift.
+Increased low freq allels can also be a signal of negative directional selection or a selective sweep, and an increase in high-frequency variants can be a signal of positive directional selection \cite{?}.
 
-There is a reduction of Ne detected in a high level of linkage of the SBPA, the lower nucleotide diversity ($\pi$ ) There is also an accumulation of rare alleles and a lot of alleles fixed in the population.
-These observations suggest that the bottleneck event and the ensuing intensification of genetic drift had a noticeable impact on the genetic composition of the SBP population.
-
-Earlier studies have shown a halving in nucleotide diversity when comparing wild soybean to their domesticated progenitors and a further halving in nucleotide diversity in breeding regimes in recent years comparing landraces to elite cultivars (ref).
-The SBPA have a mean nucleotide diversity of ($\pi$ = 0.000999), 0.09\% or a difference of  9/10.000 bases between two sequences.
-Whereas the CCA has almost double the mean nucleotide diversity of  ($\pi$ = 0.001911)  0.19\% or 19/10.000 bases differing.
-The CCA has two times higher nucleotide diversity than the SBPA.
-This is consistent with a bottleneck during the selection of founders and subsequent selection by the breeding program.
-
-The nonrandom association of alleles at different sites due to recombination (LD) in soybean shows like in the CCA
-In this comparison of CCA and SBPA, we see a high LD at both low and high ranges of the SBPA.
-The LD slope of the SBPA shows a higher degree of decay over shorter physical distances, indicating in relation to the CCA, a contrasting population size.
-
-Longer range association estimates a small Ne, because of the bottleneck. 
-The long-range LD will be due to a high level of fixation caused by drift.
-
-figure text:  The LD plot shows the relationship between genetic markers in the two populations at different sites due to recombination.
-The slope represents the rate at which LD decays with physical distance.
-The top cyan line is SBPA and the red line is the CCA.
-
-A difference in slope can also indicate different selective pressures or adaptive processes, such as the steep and higher slope of the SBPA may suggest positive selection or recent selective sweeps.
-
-AF 
-The fixation of alleles due to drift is also seen in the allele frequency of the SBPA. The high number of fixated alleles seen as 
-in a high number of allele frequencies in the extremities of the frequency spectrum is a pattern of fixation due to drift.
-Increased low freq can also be a signal of negative directional selection or a selective sweep, and an increase in high-frequency variants can be a signal of positive directional selection \cite{}.
-
-\textbf{Allele Frequency shows a high level of drift}
-The high number of alleles at high frequency shows that the SBPA have a higher number of monomorphic locus than the CCA.
-This high number of fixed alleles arises through natural selection, genetic bottlenecks, or genetic isolation in the population.
-The SBPA also has a large number of rare alleles, a high number of genetic variants in low frequencies.
-Rare alleles arise through mutation, genetic drift, or migration from other populations.
+\subsubsection{LD}
+Figure \ref{fig:LD} shows a difference in LD over distance of the combined chromosomes. The decay of LD as measured by $r^2$ in the SBPA population and the CCA population.
 
@@
 or
+
 In conjuncture, the high number of rare alleles and high number of fixed alleles together reveal the bottleneck as also seen in the following LD decay analysis.
@@
 
-\subsection{H3 Genome-wide selection signatures}
+The nonrandom association of alleles at different sites due to recombination (LD) in soybean shows like in the CCA
+In this comparison of CCA and SBPA, we see a high LD at both low and high ranges of the SBPA.
+The LD slope of the SBPA shows a higher degree of decay over shorter physical distances, indicating in relation to the CCA, a contrasting population size.
 
-Selection happened during the Swedish breeding program.
-Selection during the Swedish breeding program occurred and can be detected as the accessions are insensitive to photoperiod, and selection occurred signatures are detected at other genomic regions related to seed coat and hilum colour, cold tolerance, and flowering.
+Longer range association estimates a small Ne, because of the bottleneck. 
+The long-range LD will be due to a high level of fixation caused by drift.
 
-Selection at genes of interest.
- Selection can have a similar effect on LD in the genome as a bottleneck can, but it is a localised phenomenon in and around the targeted loci.
-We know the breeding program goals and have selected areas to target looking for selection signatures by using areas of the genome that are connected or hypothesised to be relevant to the expression involved to named traits.
-xpclr results.
+A difference in slope can also indicate different selective pressures or adaptive processes, such as the steep and higher slope of the SBPA may suggest positive selection or recent selective sweeps.
 
-haplotype groups?
+AF
 
-The Swedish soybeans are mainly a snapshot of a breeding program stopped around 1978 and all lines or recent crosses from this program were then set into the
+\begin{figure}[p]
+\centering
+\includegraphics[width=\linewidth]{plot_AF.wgs.pdf}
 
-\section{Figures and tables}
+\caption{Allele frequency distribition of the populations SBPA and CCA.  based on 8m snps with 1percent MAF filtered out. Comparing the two distrubutions }
+\label{fig:AF}
+\end{figure}
+\begin{figure}[p]
+\centering
+\includegraphics[width=\linewidth]{plot_LD_light.pdf}
 
-\subsection{PCA1 figure}
+\caption{LD decay ($r^2$) plotted against physical distance of the SNPs of the wgs data of aprox.
+8 million SNPs.
+The mean value of $r^2$ was calculated in 10kb bins.
+The LD plot shows the relationship between genetic markers in the two populations at different sites due to recombination.
+The slope represents the rate at which LD decays with physical distance.
+The top cyan line is SBPA and the red line is the CCA.
+(If there is time i will include the slopes as vanlculated by the regression model)}
 
-Figure \ref{fig:pca} shows an example figure.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{plot_PCA_source.pdf}
-\caption{Biplot of the principal component analysis of soybean genotypes.
-The colours represent the defined groups; Core collection Accessions (CCA) in black, the Swedish Breeding Program Accessions (SBPA) in green and the Founders of the breeding program (F) in red.}%
-\label{fig:pca}
+\label{fig:LD}
 \end{figure}
 
-\subsection{PCA2 figure}
+\subsection{H3 Genome-wide selection signatures of vital pathways found to significant}
+
+Two methods of detecting selection in the genome was used.
+Tajimas D score was calculated on the SBPA population at 25kbp windows, detecting the non random values by  the measured and the expected values.
+Values under 0 are indicative of selection and this was the threshold used to test for a significant change in pathways for windows with genes where the amount was significanlty different from what was expected.
+using a fisheres exact test.
+Three pathways with a total of 443 genes in these pathways were shown to have been sigificant for degredated in the population; The Phospholipases pathway (lipasyn-pwy) and the Triacylglycerol Degradation pathway (lipasyn-pwy) are both involved in lipid metobilism.
+Additionally the Aerobic Respiration I (Cytochrome C) (pwy-3781) which is an Mitocondria Energy pathway was significant.
+
+This is a signal that the degredation of these pathways has bean selected for.
+The Fst analysis showed no significants at these pathways but the odds ratio was towards degredation for the Triacylglycerol Degradation and Aerobic Respiration I (Cytochrome C) pathways.
+It is curious that pathways vital for respiration (such as Aerobic Respiration I involving Cytochrome C) and those involved in fatty acid and lipid biosynthesis and degradation (involving Phospholipases and Triacylglycerol Degradation) appear to have been selected against.
+
+Fst scores calculated by a pairwise basepair comparison of the SBPA to the CCA was used with a threshold chosen at the topå 5\%.
+What we have discovered are colocalizations of regions exhibiting elevated Fst.
+Within these regions, we have conducted tests to provide evidence of recent sweeps.
+There are 34 pathways with a total of  where there is clear signal of gene enrichment.
+Interestingly these are essential plant metabolic pathways including primary metabolism where the differences are detected.
+See table Selection for the pathways and supplementary material for the genes in these pathways.
+For this analysis, all 34 pathways indicated enrichment.
+These pathways include; Amino Acid Metabolism; Carbohydrate Metabolism; Lipid Metabolism; Metabolism and Energy Production; Nucleotide Metabolism; Secondary Metabolism and Defense; And Vitamin and Cofactor Metabolism.
+see supplemetery matheral for the full overview of pathways and genes involved.
+
+\section{Discussion}
+
+By combining population genetic and genomic methods the characterisation of the SBPA has shown that the selective breeding effort including founder selection has caused changes in the SBPA population.
+These changes are seen in light of the bottleneck effectuated on the population through the breeding program.
+We can infer that this bottleneck caused a decline in Ne and that drift acted strongly on the population, causing allele fixation.
+The selective breeding effort caused signatures of selection detected in this analysis, and its impact was observed primarily in key genetic pathways.
+
+\textbf{Conclusion and Discussion H1} 
+The evidence indicates a population stratification for the SBPA as a subpopulation and it is clear that the SBPA are a cohesive group in agreement with the IBS analysis clustering.
+The Fst analysis provides clarity on population stratification, revealing the SBPA as a distinct and identifiable subpopulation due to the high value of the fst.
+The PCA shows the SBPA grouped together due to their similarity.
+The IBS calculation show the SBPA to be similar to each other and distinct from the CCA as a whole.
+
+It is expected that the SBPA would be seen as a subpopulation within the overall population using the CCA as a proxy of this.
+Comparing these results to previous work.
+Generally, a breeding program begins with a limited number of founders; these founders may have their own unique genetic backgrounds, and their genes will become prominent in the population over time.
+Crossing in the generations of the breeding program will function as inbreeding and diversity in the subpopulation will decrease over time.
+This has been seen in papers investigating both domestication and more recent breeding trajectories.
+What have others seen?
+problems or shortcomings?
+It was important for the further analyses of this research so as to avoid a confirmation bias loop that the SBPA was designated as a distinct subpopulation.
+This substantiated claim would then need to be supported by the PCA, IBS and Fst analysis to avoid circular reasoning in the next analyses.
+
+\textbf{H2 genetic diversity decline}
+
+There is lower  genetic difference within the population of the 160 SBPA, as seen by the measure of nucleotide diversity.
+Essentially, this result indicates a high amount of genetic redundancy within the individuals of the population.
+This result is expected and follows the consequences of a breeding program where few founders are used, and the same founders (or offspring of crosses between these same founders) are continually recrossed into the breeding lines, effectuating inbreeding.
+This high similarity within the population was observed in the PCA and IBS analysis, the pi calculation confirms this and quantifies it.
+The explanation for the low nucleotide diversity is due to the initial diversity in the founders, the random drift caused by the small population size, and the common soybean breeding practices.
+Which are hybridisation to create variability for the breeder's selection followed by backcrossing and single seed decent.
+The practice of selfing the plants for several generations ensures homozygousity needed for seed suitable for cultivar distribution but further diminishes any diversity that was heterozygous \cite{acquaah12}.
+The assertion regarding the utilization of these breeding practices gains additional support from the pedigrees and historical documents detailing the breeding program and its outcomes, as presented in supplementary material S3 and referenced in  \cite{holmberg1973}.
+
+comparing these results to previous work, 
+The CCA and them In relation to the GrinGlobal. USDA
+
+Discuss any problems or shortcomings encountered during the course of the work.
+If i were to repeat this analysis i would have seen if there were any outgroups to comapre with.
+An  intersect was conducted of the 50K snp data and the wgs data to include founders for the PCA analysis , I could have included some glycine Sawyer and other glycine Max accessions, so as to have a better picture of the change in nucleotide diversity within the SBPA.
+It is also later come to light through investigation of historical documents and Gene Banks that there is a greater number of founders in the USD a 50 K snip datafuture analysis could include this larger number of founders so as to see the change from soybean diversity as a whole to the end nucleotide diversity change in the founders to the nucleotide genetic diversity change in the SBPE population.
+We are here by comparing the CCA to the SBPA debottleneck, affect or founder effect of the founders is in the data.
+
+\textbf{discussions \& conclusion H2 Bottleneck and Ne}
+summarizing the results
+High number of fixed and alleles in the SBPA population is s due to initial nucleotide diversity level and drift.
+
+analysis are well-supported and indicate certain characteristics of the SBPA population substantiated in the population assessment as a whole, this comprehensive examination of various aspects related to the SBPA as a population shown from several facets (name them) a population that has indeed gone through a bottleneck effectuating a small population size causing strong drift effects.
+
+LD
+Linkage Disequalibrium is higher in the SBP a population than the CCA.
+The CCA linkage pattern is as in other soy diversity panels \cite{liu16}.
+The rate of decay or the slope of linkage seen can be directly compared can be a direct indication of effective population size is.
+It is therefore clear in figure \ref{fig:LD} that the SBPA population has a smaller effective population size than the CCA.
+Genetic linkage is seen to increase the amount of genetic drift near any given locus thus reducing the effective population size.
+It is expected that a breeding screen would have the effect on a population such as the SBPA of higher correlation of alleles  at different loci than the CCA.
+Due to the limited number of recombination since the initial crosses of founders were created.
+Genetic drift will also create LD because of the random haplotypes drifting up in frequency.
+LD is related to effective population size by the fact that recombination rate and the value of the correlation coeeficient between allele counts $r^2$.
+The Linkage disequalibrium ($r^2$) that is calculated from the data and plotted can be fitted to a nonlinear regression model using least squares.
+$y_i=1/\alpha+\beta  c_i)  +e_i$
+$slope = \alpha$
+
+A method for inference of effective populationsize is:
+$N_e = 1 / (2 * c * E(r^2) - 1)$
+$c= recombination rate $
+the LD can be replaced by the slope: 
+$E(r^2) =\alpha$
+$Ne = 1 / (2 * \alpha * c)$
+
+These estimations are only valid under idealised conditions, but as the CCA and the SBPA both deviate from the conditions and one could speculate that they deviate in similar ways as they are of the same domesticated species undergoing similar human caused artificial selection.
+Therefor i conclude that slope and therefore the Ne of the two populations can be compaired.
+We can therefore conclude that the Ne of the SBPA is smaller that the CCA.
+
+AF
+Drift is strong in a population with small Ne because effective population size is a statistical measure of how meny copies of a locus are effectivly segregating in the population, so
+
+We know that
+
+\textbf{discussions \& conclusion H3 Selection}
+summarizing the results
+We detected 34 different pathways, varying in primary importance to plant function, with a significant number of genes involved in selective sweeps.
+Additionally, we identified significance indicating selection against three additional pathways.
+
+was this the expected result? No, not really, once it was clear that the Ne of the SBPA was small and drift, having a strong influence on the population that there would not be much signal detected. 
+But other studies combining the population genomics (they did the fst sweeps similar to me) did also find evedence of lokal adaptation in aribidopsis \cite{price18}in soybean \cite{FIND} 
+Do we see evidence of selection? 
+yes
+Are we looking in the right places?
+Should we expect directional changes for these traits? or are part of these traits complex  traits
+
+Discuss signal detection
+Signal. Population
+Ne
+selection pressure
+
+Discuss the data
+sample size
+time of the breeding program?
+
+Discuss the method and whether or not it is able to detect what we want to know. 
+Analysis choice
+Threshold arbitrary
+caveats
+
+Conclusion and Implications
+this study reports the characteristics of the SBPA population  
+ramifications and such
 
@@
 
-\subsection{Dendogram figure}
-
-Figure \ref{fig:dendo} shows an example figure.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{rainbow.pdf}
-\caption{Hierarchical cluster dendrogram based on pairwise identity-by-state (IBS) values from SNP data for all samples. describe dendogram}%
-\label{fig:pca}
-\end{figure}
-
-\subsection{Heatmap figure}
-
-Figure \ref{fig:heat} .
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{plot_heatmap.pdf}
-\caption{PCA regions}
-\label{fig:heat}
-\end{figure}
-
-\subsection{Table 1}
-
-Table  \ref{tab:fst} shows table 1.
-
-\section{other text}
-
-abstract:
-Soybean, belonging to the Fabaceae family and scientifically known as \textit{Glycine max} (L.
-Merrill), is globally recognised as the primary legume crop due to its high protein and oil content.
-However, its genetic diversity has decreased over time due to domestication and intensive artificial selection, driving the need to look elsewhere to increase germplasm utilisation for abiotic adaptation \citep{hyten06, gizlice96}.
-Relevant accessions with the potential for abiotic adaptation to cold tolerance have been identified [@haupt20].
-Additionally 160 Swedish accessions, originating from a Swedish breed for the adaptation to the cool climate of north western Europe and stored in the NordGen gene bank have been unearthed [@holmberg1973].
-We will interpret the plant genetic resources available, so as to utilise the genetic potential.
-To facilitate targeted selection of promising accessions from soybeans vast germplasm collections,  and unveiling the genetic data stored in the NordGen seed banks.
-findings are: ...
-
-\textbf{Founders}
-Founders identified through pedigrees and other genebank metadata have differing similarities to the SBPA.
-
-\textbf{Ugra soja}
-Ugra soja 
-"Svalöf’s Ugra soybean is a new variety obtained from a cross between a black-seeded American variety, Wisconsin, and a brown-seeded Polish variety, Brunatna Vilenska. " soybean history book
-
-'Ugra' 
-labeled "Ugra", Ugra Saja" & "Ugra Soja"
-Developed by Svalöf AB in Malmöhus, Sweden.
-Pedigree (Wisconsin Black' with 'Brunatna Vilenska' (a Polish variety).)  It was released in 1950 (source:nordgen).
-USDA accession number "Ugra Saja"  is PI 198067 and "Ugra Soja"  at Nordgen NGB 23807
-
-Notes on svalof:
-
-    Svalöf AB later merged with Weibulls AB to form Lantmännen SW seed, a partly Swedish owned seed company.
-The company is headquartered in Landskrona (Southern Sweden).
-The town Landskrona is located within the county administration area called Malmö (once called Malmöhus).
-    "‘Ugra saja’ is actually just named Ugra.
-The ‘saja’ is probably just a misspelling of soja (soya in Swedish)." - Dr.
-Fredrik Fogelberg, Swedish Institute of Agricultural and Environmental Engineering."
-
-\textbf{Altongård}
-
-?
-Holmberg has been generous in sharing his germplasm with other soybean breeders.
-Several Canadian workers have used his lines in their crosses.
-For example, one of the parents of ‘Altona’ is a Holmberg line.?
-
-other text
-
-using the 8 million SNPs of the 409 CCA and 153 SBPA soybeans and the .
-
-To better understand the population relationships of the accessions received from the Nordic gene bank and how they are related to each other, and the rest of the soybean germplasm, they are here compared to a Core Collection of 409 soybeans from the USDA genebank chosen/curated from the soybean germplasm for Central European breeding.
-The Core collection used here is a subset of the soybean germplasm collection, the  USDA genebank accessions, which are assessed to be  preadapted to cultivation in Central Europe, while conserving a high genetic diversity \cite{haupt20}.
-Additionally, accessions were grouped into categories based on the suspected breeding history.
-There are the ten suspected Founders of the Swedish breeding program and the Swedish accessions of the Nordgen genebank.
-
-A Principal Component Analysis (PCA) and the genome-wide Identity By State (IBS) pairwise distance matrix were applied to visualise the overall variance of the data.
-The resulting PCA (Figure \ref{fig:pca}) shows the clustering of the Swedish accessions as a sub-group based on similarity.
-The variance proportions explained.
-
-A closer look at the PCA revealed two accessions of the Nordgen Swedish that were not part of the particular breeding program of interest as seen in (Figure \ref{fig:pca3}  showing the PCA indicating the two accessions not from the breeding program in the cluster with the CCA in orange and black)  and also thereafter confirmed by the genebank passport data that they originated from other Swedish soybean breeding programs (mention more here when certain.
-Altonagård and Ugra soja/ saja, malmåhus).
-This figure also shows the accessions of the Nordgen and core collection that are Swedish of origin and with a name consisting of letters (excluding numbers or non-alphanumeric characters of the accessions' names) as included in the database.
-The IBS (Figure  \ref{fig:dendo}) clearly shows the Swedish accessions of the breeding program of interest clustering together. The  
- IBS numbers in a table showing ibs within sbpa and out?
-
- Heatmap based on the sbpa shows high similarity of the SBPA and shows that the 2 accessions Altonagård/Altonagard and Ugra Soja(Saja) are dissimilar.
-
-The population differentiation measured by $F_{\text{ST}}$
-i think we agreed that we needed
-
-change values:
-Fst estimation on genotypes:
-Excluding 0 SNP (monomorphic: TRUE, MAF: NaN, missing rate: NaN)
-    # of samples: 564
-    # of SNPs: 8,533,444
-Method: Weir & Hill, 2002
-### of Populations: 2
-    CCA (409), SBPA (155)
-[1] 0.2412619
-[1] 0.1783931
-     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
--0.002281  0.048690  0.136954  0.178393  0.263807  0.938701 
 \begin{table}[p]
@@
 
-\begin{figure}
+\begin{table}[tb]
 \centering 
-    \includegraphics[width=0.5\linewidth]{Screenshot 2023-07-02 at 14.40.39.png}
-    \caption{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3517337/}
-    \label{fig:enter-label}
-\end{figure}
-
-\begin{table}[p]
-\centering 
-\caption{$F_{\text{ST}}$ estimation on Maturity groups, measuring the population differentiation of  the 155 Accessions of the Nordgen genebank that were part of the breeding program)(SBPA) to the Core Collection of 409 accessions (CCA) groups by the corosponding maturist groups.}
+\caption{$F_{\text{ST}}$ estimation on Maturity groups, measuring the population differentiation of  the 155 Accessions of the Nordgen genebank that were part of the breeding program)(SBPA) to the Core Collection of 409 accessions (CCA) groups by the corrosponding maturity groups.}
 
@@
 
-postulate: Population subdivision analysis will quantify the SPBA will be a distinct subpopulation with a moderate to distinct allele frequency difference.
-
 \section{Supplementary Material}
@@
 
-## 
-## To cite R in publications use:
-## 
-##   R Core Team (2021). R: A language and environment for statistical
-##   computing. R Foundation for Statistical Computing, Vienna, Austria.
-##   URL https://www.R-project.org/.
-## 
-## A BibTeX entry for LaTeX users is
-## 
-##   @Manual{,
-##     title = {R: A Language and Environment for Statistical Computing},
-##     author = {{R Core Team}},
-##     organization = {R Foundation for Statistical Computing},
-##     address = {Vienna, Austria},
-##     year = {2021},
-##     url = {https://www.R-project.org/},
-##   }
-## 
-## We have invested a lot of time and effort in creating R, please cite it
-## when using it for data analysis. See also 'citation("pkgname")' for
-## citing R packages.
+add to file 
+   @Manual{,
+     title = {R: A Language and Environment for Statistical Computing},
+     author = {{R Core Team}},
+     organization = {R Foundation for Statistical Computing},
+     address = {Vienna, Austria},
+     year = {2021},
+     url = {https://www.R-project.org/},
+     }
 
